### 2023-05-03 manuscript/intro.qmd

# Introduction

Rp smooth, dynamics work
-> [0,1], do the dynamics work ?
there are non-continuous embeddings, so neural field disc.
is not continuous -> does it break the dynamics ?
                  -> are dynamics well-defined in [0,1] ?
                     [cite Spatiotemporal dynamics of continuum neural fields : in litterature
					   the kernel is always continuous. ]
				  -> here the "field" is nowhere continuous (fractal), but is it still well-defined
				  -> numerical answers

why low-rank ? -> it is easy to simulate, we just use it to illustrate our examples @schmutz2023convergence

do not ask this question "can we write a neural field ?" 
but rather this       -> "what neural field can we write given a regularity ?"

here we have the first example of two equivalent dynamics. one with high regularity and dimension.,
the other with low regularity and dim.

why are we doing this ?
dimensionality of activity = rank(connectivity)
normally, we would think that if we can write a neural field equation, then the embedding space if of dimension rank(connectivity).
-> but we have shown that is not the case ! with the correct mapping, we can write a neural field in 1D, even though the dimensionality of the activity is 2D !
-> not even a tradeoff. IDEA : smooth function in 1D. when we map back to 2D using recusive local, the mapped space is everywhere discontinuous, but (TEST THIS !!) the dynamics are the same. -> THE DIMENSION OF THE EMBEDDING IS UNRELATED TO THE DIMENSION OF THE ACTIVITY. Rather, we should ask in which dimension/embedding the connectivity kernel is going to have a nice expression (e.g. continuous etc)

@pham2016existence

![connectivity](figures/connectivity_binning.png)