### 2023-05-08 manuscript/chapters/3_dynamics_1d.qmd

# Mapping $\mathbb R^2$ to $[0,1]$ and resulting dynamics in $[0,1]$ {#sec-mapping}

## Which types of mappings

peano vs cantor mappings

linear projections don't work ofc

## Sequence of mappings and their limit

intuition for reshape, and the fractals

expose different types of mapping

## Numerically implementing mappings to $[0,1]$

how to we actually simulate the mappings in one-dimension

go into detail on how numerically S is a reordering of the bins

## Simulating the $[0,1]$ neural field

how we numerically implemement a simulation on a fractal

averaging of 2^n neurons