### 2023-05-08 manuscript/index.qmd

# Index {.unnumbered .hidden}

\[TODO : a nice illustration for this, maybe some cool animation ?\]