### 2023-05-31 manuscript/chapters/4_locality.qmd
yay for more writing !

# Locality and numerical convergence

## Analytical equivalence of the dynamics

composition of functions

bijectivity (actually injection in one sense)

## Locality

the notion of infinite number of discontinuities, but the size counts too

plot distance 2D vs distance 1D
* "close to the line y=sqrt2 x" (but fractals dont get arbitrarily close)
* "size of the discontinuities"
* L1 distance

“resilient to some discontinuity”
the notion of infinite number of discontinuities, but the size counts too

Total variation vs bining variation

histogram of discontinuities ?

$$
\frac{1}{2^n}\sum_{i=1}^{2^n} \sup_{z,z' \in \{\frac{k}{4^n} | k= 1, \cdots, 2^n\}} |f^n_1(\tfrac{i-1}{2^n}+z) - f^n_1(\tfrac{i-1}{2^n}+z')| + |f^n_2(\tfrac{i-1}{2^n}+z) - f^n_2(\tfrac{i-1}{2^n}+z')|
$$

plot the bins in the 2D scatter, so we can see what we are doing.

it is different to TV, TV is projecting all the jumps

## Numerical feasibility

failed notions of discontinuity

locality might be a side-effect from bijections

measure for V

proof that numerical convergence

## Details

ironing out details : here we considered 2d -> 1d, but possible to repeat the application of the mapping so that Rp -> Rp-1 -> .. -> 1d

also should work with any neural field, here low-rank cycling was just for the purpose of demonstration