### 2023-06-04 manuscript/chapters/4_locality.qmd
this takes so much time nyaaa

::: {.content-hidden}
{{< include ../macros.tex >}}
:::

# Locality and numerical convergence

## Analytical equivalence of the dynamics

composition of functions

bijectivity (actually injection in one sense)

TODO : write the proof of this

## Locality

### Intuition behind the notion of locality

![](figures/distance_2d_1d.png)

localité :
1. proche in 2D -> proche in 1D (OK projection (obv., x est proche), OK zorder, PAS random)
2. proche in 1D -> proche in 2D (PAS projection, PAS random, OK zorder)

[TODO EXPLAIN : explain this better. basically the thing is we AVERAGE in 1D, so this better be meaningful]

the notion of infinite number of discontinuities, but the size counts too

plot distance 2D vs distance 1D
* "close to the line y=sqrt2 x" (but fractals dont get arbitrarily close)
* "size of the discontinuities"
* L1 distance

"resilient to some discontinuity" the notion of infinite number of discontinuities, but the size counts too

![histogram of discontinuities](figures/image_20230517_2.png)

### The total variation attempt

Let $S^{-1} : \mathbb R \rightarrow \mathbb R^2$, let $|| \cdot ||$ be a norm in $\mathbb R^2$ (any norm is OK, since all equivalent (i.e. equal to a constant)). Here we use the $\ell^1$ norm (absolute value).

$$
\begin{aligned}
\mathrm{TV}(S^{-1}) &= \sup_{P \in \mathcal P} \sum_{i=1}^{n_P} || S^{-1}(z_{i}) - S^{-1}(z_{i-1}) ||_1 \\
&= \sup_{P \in \mathcal P} \sum_{i=1}^{n_P} | S^{-1}_1(z_i) - S^{-1}_1(z_{i-1})| + | f_2(z_i) - f_2(z_{i-1}) |
\end{aligned}
$$

this basically just projects the discontinuities on x and y.

in a way since you need to go to each point only once, so there is only so much you can do inside the $[0,1]^2$. the tiny jumps still add up, and all mappings seem to grow $TV$ in the same way

![](figures/image_20230517.png)

-> what we really are expression, is the fact that doing the averaging in 1D is not a big problem when the mapping is good.

### Mathematical formulation of average binned variation

variation inside the bins, weak version

$$
\begin{aligned}
V_n(S^{-1}) &= \frac{1}{2^n}\sum_{i=1}^{2^n} \sup_{\delta,\delta' \in [0, \tfrac{1}{2^n}]^2} |S_1(\tfrac{i-1}{2^n}+\delta) - S_1(\tfrac{i-1}{2^n}+\delta')| + |S_2(\tfrac{i-1}{2^n}+\delta) - S_2(\tfrac{i-1}{2^n}+\delta')| \\
&= \frac{1}{2^n}\sum_{i=1}^{2^n} \sup_{\alpha, \alpha^\prime \in [\tfrac{i-1}{2^n}, \tfrac{i}{2^n}]^2} \norm{S^{-1}(\alpha^\prime) - S^{-1}(\alpha)}_1
\end{aligned}
$$

variation inside the bins, strong version. basically we take the max over all binning methods that have bins smaller than $\delta$. Let $P_\delta$ be a permutation of $[0,1]$ with $n_{P_\delta}$ elements $\{\alpha_1 \leq \alpha_2 \leq \cdots \alpha_{n_{P_\delta}}\}$, where $\alpha_{i}-\alpha_{i-1} \leq \delta$, and $\mathcal{P}_\delta$ the set of such permutations.

$$
\begin{aligned}
V_\delta(S^{-1}) &= \sup_{P_\delta \in \mathcal{P}_\delta} \delta \sum_{i=1}^{n_{P_\delta}} \sup_{\alpha, \alpha^\prime \in [\alpha_{i-1}, \alpha_{i}]^2} \norm{S^{-1}(\alpha^\prime) - S^{-1}(\alpha)}_1 \\
V &= \limsup_{\delta \to 0} V_\delta
\end{aligned}
$$

evidently,

$$
0 \leq V_n \leq V_\delta
$$

[TODO FIG : show the range of each bin, how each bin contains $2^n$ smaller bins. plot the bins in the 2D scatter, so we can see what we are doing]

[TODO MATH : the exact norm is not important. all norms are equivalent]

[TODO NOTE : in the definition of $V$, we use the converged version of $S$, instead of a finite truncation $S^n$. we need to convince people this is not a big mistake, since $S^n \to S$, and we can take both limits (binning and $S^n$) "at the same time"]

### Average variation of mappings

Numerical demonstration

![](figures/v_num.png)

In order to compute (a bound on) $V_n$ for some mappings, we consider the binary expansion of $\alpha$ and $\alpha^\prime$ which appear in the supremum. For clarity and to represent actual numerical implementations, we furthermore truncate the binary expansion to $2m$ bits (on most machines, $2m=64$ bits).

$$
\begin{aligned}
\alpha &= 0. b_1 b_2 \cdots b_{2m-1} b_{2m}\\
\alpha^\prime &= 0. b_1^\prime b_2^\prime \cdots b_{2m-1}^\prime b_{2m}^\prime \\
\end{aligned}
$$

Furthermore, since $\alpha \in \left[\frac{i-1}{2^n}, \frac{i}{2^n} \right],\, i \in \{1, \cdots, 2^n\}$, we may express the first $n$ (where $n < 2m$) bits of $\alpha$ with the bits of $i=\sum_{k=0}^{n-1} i_k 2^k$ (we ignore $i=2^n$ for the sake of simplicity, which would introduce one more bit, $n+1$ in total).

$$
\begin{aligned}
\alpha &= 0. i_{n-1} i_{n-2} \cdots i_1 i_0 b_n b_{n+1} \cdots b_{2m-1} b_{2m}\\
\alpha^\prime &= 0. i_{n-1} i_{n-2} \cdots i_1 i_0 b^\prime_n b^\prime_{n+1} \cdots b^\prime_{2m-1} b^\prime_{2m}
\end{aligned}
$$

#### Z-mapping

Let us consider the Z-mapping, and the two terms compositing its inverse. In the case of $\alpha, \alpha^\prime$ above, we have if $n$ is even :

$$
\begin{aligned}
Z^{-1}_1(\alpha) &= 0.\underbrace{i_{n-1} i_{n-3}}_\text{$n/2$ bits} \cdots i_1 b_n b_{n+2} \cdots b_{2m-1} \\
Z^{-1}_2(\alpha) &= 0.\underbrace{i_{n-2} i_{n-4}}_\text{$n/2$ bits} \cdots i_0 b_{n+1} b_{n+3} \cdots b_{2m}
\end{aligned}
$$

If $n$ is odd, we have :

$$
\begin{aligned}
Z^{-1}_1(\alpha) &= 0.\underbrace{i_{n-1} i_{n-3}}_\text{$(n+1)/2$ bits} \cdots i_0 b_n b_{n+2} \cdots b_{2m-1} \\
Z^{-1}_2(\alpha) &= 0.\underbrace{i_{n-2} i_{n-4}}_\text{$(n-1)/2$ bits} \cdots i_1 b_{n+1} b_{n+3} \cdots b_{2m}
\end{aligned}
$$

We only consider even $n$ in the following, as the proof remains very similar.

The core of the proof relies on the fact the first $n/2$ bits of the difference of the inverses cancel out when computing the average binned variation.

$$
\begin{aligned}
|Z^{-1}_1(\alpha) - Z^{-1}_1(\alpha^\prime)| &= 0.0 \cdots 0 (b_n b_{n+2} \cdots b_{2m-1} - b^\prime_n b^\prime_{n+2} \cdots b^\prime_{2m-1}) \\
&= 2^{-n/2} \sum_{k=1}^{\tfrac{2m-n+1}{2}} (b_{n+2k-1} - b^\prime_{n+2k-1}) 2^{-k} \\
&\leq 2^{-n/2} \\
|Z^{-1}_2(\alpha) - Z^{-1}_2(\alpha^\prime)| &= 2^{-n/2} \sum_{k=1}^{\tfrac{2m-n+1}{2}} (b_{n+2k} - b^\prime_{n+2k}) 2^{-k} \\
&\leq 2^{-n/2}
\end{aligned}
$$

Putting everything together, we get an upper bound for the average binned variation.

$$
V_n(Z^{-1}) \leq \frac{1}{2^n} \sum_{i=1}^{2^n} 2^{-n/2} + 2^{-n/2} = 2^{-n/2+1} \xrightarrow{n \to \infty} 0
$$

#### Column mapping

We now consider the column mapping. Since here we consider finite-precision $\alpha$ and $\alpha^\prime$, the invere is well-defined and we have

$$
\begin{aligned}
C^{-1}_1(\alpha) &= 0.i_{n-1} i_{n-2} \cdots i_1 i_0 b_n b_{n+1} \cdots b_m \\
C^{-1}_2(\alpha) &= 0.b_{m+1} b_{m+2} \cdots b_{2m}
\end{aligned}
$$

The first component can be bounded using the same trick from the previous proof :

$$
\begin{aligned}
|C^{-1}_1(\alpha) - C^{-1}_1(\alpha^\prime)| &= 2^{-n} \sum_{k=1}^{m-n+1} (b_{n+k-1} - b^\prime_{n+k-1}) 2^{-k} \\
&\leq 2^{-n}
\end{aligned}
$$

However now problems arise when we compute the averge binned variation, since the second component is not well-behaved. In particular, since we take the $\sup_{\alpha,\alpha^\prime \in [\tfrac{i-1}{2^n},\tfrac{i}{2^n}]^2}$, we can pick the specific values such that $b_k=1$ and $b^\prime_k=0$ for all $k \in \{m+1, \cdots 2m\}$.

$$
\begin{aligned}
|C^{-1}_2(\alpha) - C^{-1}_2(\alpha^\prime)| &= \sum_{k=m+1}^{2m} (b_k - b^\prime_k) 2^{-(k-m)} \\
&= \sum_{k=m+1}^{2m} 2^{-(k-m)} \\
&> \frac 12
\end{aligned}
$$

Putting both together, we can write :

$$
\begin{aligned}
V_n(C^{-1}) &= \frac{1}{2^n}\sum_{i=1}^{2^n} \sup_{\alpha, \alpha^\prime \in [\tfrac{i-1}{2^n}, \tfrac{i}{2^n}]^2} |C^{-1}_1(\alpha^\prime) - C_1^{-1}(\alpha)| + |C^{-1}_2(\alpha^\prime) - C_2^{-1}(\alpha)| \\
&\geq \frac{1}{2^n}\sum_{i=1}^{2^n} \sup_{\alpha, \alpha^\prime \in [\tfrac{i-1}{2^n}, \tfrac{i}{2^n}]^2} |C^{-1}_1(\alpha^\prime) - C_1^{-1}(\alpha)| + \sup_{\alpha, \alpha^\prime \in [\tfrac{i-1}{2^n}, \tfrac{i}{2^n}]^2} |C^{-1}_2(\alpha^\prime) - C_2^{-1}(\alpha)| \\
&\geq \frac 12 + \bO(2^{-n})
\end{aligned}
$$

Therefore we have found a lower bound, which proves the average binned variation does not converge to zero (if it converges at all).

$$
V_n(C^{-1}) > \frac 12 \implies \lim_{n \to \infty} V_n > \frac 12
$$

## Numerical feasibility

numerically, we have finitely many bits, and cannot fully represent the fractal. we therefore need the finite approximations to be already good.

locality might be a side-effect from bijections

Before we make the proof of numerical feasibility.

In the evaluation of the evolution equation, the integral term in each coordinate $\alpha$ is problematic.

Let us define (recall) the numerical and analytical integral (we leave a $w_U(\cdot, \beta)$ as implicit function of the first argument $\alpha$). The intuition behind this proof of convergence is that the variation inside of each numerical bin vanishes as we increase the number of iterations $n$ (even though the structure becomes fractal !). The variation comes from the coarse-graining, where we go from $4^n$ to $2^n$ samples.

We define $\tilde h_i(t) = \tilde h(t, \tfrac{i-1}{2^n}) = \tilde h(t, \beta_i),\,\beta_i=\tfrac{i-1}{2^n},\,i\in\{1,\cdots,2^n\}$.

$$
\begin{aligned}
\mathrm{NI}_n &= \frac{1}{2^n} \sum_{i=1}^{2^n} \tilde w(\cdot, \beta_i) \phi(\tilde h_i) \\
\mathrm{AI} &= \int_{[0,1]} \tilde w(\cdot, \beta) \phi(\tilde h(t, \beta)) \lambda \circ S^{-1}(\d \beta)
\end{aligned}
$$

We wish to prove that $|\mathrm{NI}_n - \mathrm{AI}| \xrightarrow{n \to \infty} 0$ (implicitly : this is valid for all $\alpha \in [0,1]$).

The first part of the proof involves splitting the integral on $[0,1]$ into $2^n$ integral over segments of length $2^{-n}$, and using the triangle inequality.

$$
\begin{aligned}
|\mathrm{NI}_n - \mathrm{AI}| &= \left| \frac{1}{2^n} \sum_{i=1}^{2^n} \tilde w(\cdot, \beta_i) \phi(\tilde h_i) - \int_{0}^{1} \tilde w(\cdot, \beta) \phi(\tilde h(t, \beta)) \lambda \circ S^{-1}(\d \beta) \right| \\
&= \left| \frac{1}{2^n} \sum_{i=1}^{2^n} \tilde w(\cdot, \beta_i) \phi(\tilde h_i) - \sum_{i=1}^{2^n} \int_{\tfrac{i-1}{2^n}}^{\tfrac{i}{2^n}} \tilde w(\cdot, \beta) \phi(\tilde h(t, \beta)) \lambda \circ S^{-1}(\d \beta) \right| \\
&= \left| \sum_{i=1}^{2^n} \frac{1}{2^n} \tilde w(\cdot, \beta_i) \phi(\tilde h_i) - \int_{\tfrac{i-1}{2^n}}^{\tfrac{i}{2^n}} \tilde w(\cdot, \beta) \phi(\tilde h(t, \beta)) \lambda \circ S^{-1}(\d \beta) \right| \\
&\leq \sum_{i=1}^{2^n} \left| \frac{1}{2^n} \tilde w(\cdot, \beta_i) \phi(\tilde h_i) - \int_{\tfrac{i-1}{2^n}}^{\tfrac{i}{2^n}} \tilde w(\cdot, \beta) \phi(\tilde h(t, \beta)) \lambda \circ S^{-1}(\d \beta) \right| \\
\end{aligned}
$$

We now want to prove that evaluating the integrand at $\beta_i = \tfrac{i-1}{2^n}$ gives a good approximation of the integral. Since $\int_{\tfrac{i-1}{2^n}}^{\tfrac{i}{2^n}} \lambda \circ S^{-1}(\d \beta) = \tfrac{1}{2^n}$ [TODO MATH : is this true ?], we write

$$
\begin{aligned}
|\mathrm{NI}_n - \mathrm{AI}| &\leq \sum_{i=1}^{2^n} \left| \int_{\tfrac{i-1}{2^n}}^{\tfrac{i}{2^n}} \left[ \tilde w(\cdot, \beta_i) \phi(\tilde h_i) - \tilde w(\cdot, \beta) \phi(\tilde h(t, \beta)) \right] \lambda \circ S^{-1}(\d \beta) \right| \\
&= \sum_{i=1}^{2^n} \left| \int_{\tfrac{i-1}{2^n}}^{\tfrac{i}{2^n}} \left[ w_U(\cdot, S^{-1}(\beta_i)) \phi(h_U(t, S^{-1}(\beta_i))) - w_U(\cdot, S^{-1}(\beta)) \phi(h_U(t, S^{-1}(\beta))) \right] \lambda \circ S^{-1}(\d \beta) \right|
\end{aligned}
$$

The main source of variance inside the integrand is the term $S^{-1}$, since $w_U$ and $h_U$ are "nice" functions. We express this regularity by the added assumption that $w_U(\cdot, \vec z) h_U(t, \vec z)$ is Lipschitz in $\vec z$, with Lipschitz constant $L(\cdot)$, and $\norm{\cdot}$ the corresponding norm on $[0,1]^2$.

Once we have applied the Lipschitz property, we make use of our intuition of "locality" : points close in $[0,1]$ should be (on average) mapped to points close in $[0,1]^2$. Therefore, taking the sup inside each $2^{-n}$ segment does not introduce a too large error.

[TODO MATH : prove that it is indeed Lipschitz. The main problem I've enountered with this is that these functions are $\textrm{regular function} \circ \textrm{CDF}^{-1}$, and the inverse CDF goes to infinity at 0 and 1].

$$
\begin{aligned}
|\mathrm{NI}_n - \mathrm{AI}| &\leq L(\cdot) \sum_{i=1}^{2^n} \left| \int_{\tfrac{i-1}{2^n}}^{\tfrac{i}{2^n}} \norm{S^{-1}(\beta_i) - S^{-1}(\beta)} \lambda \circ S^{-1}(\d \beta) \right| \\
&\leq L(\cdot) \sum_{i=1}^{2^n} \left| \int_{\tfrac{i-1}{2^n}}^{\tfrac{i}{2^n}} \sup_{\beta \in [\tfrac{i-1}{2^n}, \tfrac{i}{2^n}]} \norm{S^{-1}(\beta_i) - S^{-1}(\beta)} \lambda \circ S^{-1}(\d \beta^\prime) \right| \\
&= L(\cdot) \frac{1}{2^n} \sum_{i=1}^{2^n} \sup_{\beta \in [\tfrac{i-1}{2^n}, \tfrac{i}{2^n}]} \norm{S^{-1}(\beta_i) - S^{-1}(\beta)} \\
&\leq L(\cdot) \frac{1}{2^n} \sum_{i=1}^{2^n} \sup_{\beta, \beta^\prime \in [\tfrac{i-1}{2^n}, \tfrac{i}{2^n}]^2} \norm{S^{-1}(\beta^\prime) - S^{-1}(\beta)} \\
&= L(\cdot) V_n \\
&\xrightarrow{n \to \infty} 0
\end{aligned}
$$

We see that it is the notion of "locality" of the mapping $S$ that allowed us to write the convergence of the numerical integral to the analytical value. In doing so, we exploited the fact that the "variance" inside each small segment in $[0,1]$ is small on average, and that the errors vanish as we take finer and finer bins.

## Generalizing to $[0,1]^p \mapsto [0,1]$

ironing out details : here we considered $[0,1]^2 \mapsto [0,1]$, but the mappings can be generalized to any number of dimensions.

Example for the Z-mapping. Let $\vec \alpha = (\alpha_1, \cdots, \alpha_p)$, and $\alpha_\mu = \sum_{k=1}^{n} b^\mu_k 2^{-k}$. Then

$$
Z^n(\alpha_1, \cdots, \alpha_p) = \sum_{k=1}^{n} \sum_{\mu=1}^p b^{\mu}_k 2^{-(p(k-1)+\mu)}
$$

also should work with any neural field, here low-rank cycling was just for the purpose of demonstration