### 2023-06-07 manuscript/index.qmd

::: {.content-hidden}
# Home {.unnumbered}
:::

[TODO : a nice illustration for this, maybe some cool animation ?]