### 2023-06-11 manuscript/appendices/B_zcurve.qmd
guess who just went through half a book in one day :))

::: content-hidden
{{< include ../macros.tex >}}
\def\C{\mathcal{C}}
:::

# On the quasi-bijectivity and continuity of the Z-mapping

We study the Z-curve $Z^{-1}$

## The Z-curve is not injective {#sec-zcurve-noninjective}

this is common misconception, and people write it is a bijection, when really it is not

The Z-curve : [0,1] -> [0,1]^2 is not injective since the binary expansion is ambiguous

x = 0.10000... = 0.011111...

=> alpha = 0.1y0y0y0  !=   alpha = 0.0y1y1y1y1

but the point on which the expansion is ambiguous form a set of zero measure, so the integral is still ok

on the set [0,1] \ {problems}, Z-curve is bijective

injectivity is the property that f(x) != f(y)  iff  x != y,  (equiv f(x) == f(y) iff x == y)

Z^{-1} defines a space-filling curve

not a problem because we are just repeating the same population twice

## $Z^{-1}$ defines a space-filling curve on a subset of $[0,1]$ {#sec-zcurve-lebesgue}

### The Cantor set $\C$

the cantor set has measure 0, but has the same cardinality of the continuum

### $Z^{-1} | \C$ is a surjection

### $Z^{-1} | \C$ is continuous

## Relation with the Lebesgue space-filling curve

we would also relax and say the Z-curve is discontinuous, because lim alpha ->0.5+ Z(alpha) != lim alpha -> 0.5- Z(alpha)
-> NO ! because we linearly interpolate between the quadrants. we say the curve is continuous, in the sense that it is connected. but it is allowed to jump around !

talk about the fact we have other space filling curves, such as hilbert curve or peano curve, that are continuous and surjective [0,1] -> [0,1]^2. we use the Z-curve because it's easy to implement numerically