### 2023-06-13 manuscript/chapters/3_mappings.qmd
nearly finished chap 3 !

# Mapping neural fields to lower-dimensional embeddings {#sec-mappings}

::: {.content-hidden}
{{< include ../macros.tex >}}

```{python}
import matplotlib as mpl, matplotlib.pyplot as plt, matplotlib.colors as mplcolors, matplotlib.patheffects as pe, matplotlib.ticker as ticker
import numpy as np
import sys
sys.path.append('../../notebooks')  # import neurodyn
from neurodyn import *
from matplotlib_tufte import *
setup()

mpl.use('svg')
```
:::

In the previous chapter, we have established a simple toy model for a neural field in $p$ dimensions, and have shown how it can be embedded in $[0,1]^p$. In this chapter, we discuss in-depth what types of mappings can be used to map the $p$-dimensional space to $[0,1]$, and in particular first study the case $p=2$. Before getting into too much detail, let us give some insight as to why this problem is not trivial.

## Population dynamics: an introductory example of a 1D embedding

### Embedding a finite number of populations in $[0,1]$

Let us for a moment consider another type of model for networks of neurons, that of homogeneous population dynamics, that describes how $N_\text{pop}$ different populations interact with each other. These types of models emerge when networks of neurons exhibit a finite number of "neuron types". We then group similar neurons ad make the approximation that all neurons inside the population have the same potential, and the same interaction weights with other neurons. In effect, every population then represents in actuality an infinity of discrete identical neurons.

The result of this is that the population dynamics can be written with the same equation as a finite network of neurons, but it already represents the $N \to \infty$ limit. We write the potential of the populations $H_a(t)$ and the population connectivity $\tilde J_{ab}$, where $a,b \in \{1, \cdots, N_\text{pop}\}^2$ in @eq-pop-dynamics.

$$
\dot H_a(t) = -H_a(t) + \sum_{b=1}^{N_\text{pop}} \tilde J_{ab} \phi(H_b(t))
$$ {#eq-pop-dynamics}

We don't show the derivation of population dynamics here (see @GerKis14, chapter 12.3 for a detailed analysis), but we point out the implicit scaling behavior of $\tilde J_{ab}$. Letting $i$ be a neuron of population $a$ and $j$ be a neuron of population $b$, the connectivity between the two neurons is $J_{ij}$ and the population connectivity scales with the number of neurons in the presynaptic population $|b|$. Furthermore, we make the reasonable assumption that $J_{ij}$ scales as $J^0_{ij} / N$, where $J^0_{ij} \sim \bO(1)$.

$$
\tilde J_{ab} = \frac{|b|}{N} J^0_{ij}
$$

As $N\to\infty$, the ratio $\tfrac{|b|}{N}$ converges to a constant $p_b$, the probability of the population, normalized such that $\sum_{a=1}^{N_\text{pop}} p_a = 1$. It expresses the probability of uniformly sampling a neuron belonging to population $b$. In effect, $p_b$ can be interpreted as the "weight" of the population.

So far, we have not considered that the populations are embedded in some abstract space; but the notion of population weights gives rise to the obvious embedding in $[0,1]$ given by the cumulative sums of the population weights, illustrated in @fig-popdynamics-01. We define (somewhat informally) the connectivity kernel $w(\alpha, \beta) = J^0_{ab}$ for $\alpha, \beta$ in the regions corresponding to populations $a$ and $b$ respectively.[^poplocation]

[^poplocation]: Formally, we would write the cumulative sum $K_a = \sum_{b=1}^a p_b$. Then, we have $\alpha \in [K_{a-1}, K_{a})$ and $\beta \in [K_{b-1}, K_{b})$. We abusively wrote $J^0_{ab}$ since all neurons in each population are identical, therefore we can index into the connectivity $J^0_{ij}$ with the population indices $a,b$ instead of neuron indices $i,j$.

![(TODO) Embedding of the population dynamics in $[0,1]$](figures/photo_2023-06-10_19-26-49.jpg){#fig-popdynamics-01}

Defining the neural field $H(\alpha, t) = H_a(t)$, we can write a "neural field equation" on the $[0,1]$ space.

$$
\begin{aligned}
\partial_t H(\alpha, t) &= -H(\alpha, t) + \int_0^1 w(\alpha, \beta) \phi(H(\beta, t)) \d \beta \\
&= -H(\alpha, t) + \sum_{b=1}^{N_\text{pop}} p_b w(\alpha, b) \phi(H_b(t)) \\
\iff \partial_t H_a(t) &= -H(a, t) + \sum_{b=1}^{N_\text{pop}} p_b w(a, b) \phi(H_b(t)) \\
&= -H(a, t) + \sum_{b=1}^{N_\text{pop}} p_b J^0_{ab} \phi(H_b(t)) \\
&= -H(a, t) + \sum_{b=1}^{N_\text{pop}} \tilde J^0_{ab} \phi(H_b(t)) 
\end{aligned}
$$

We hereby showed (having taken many writing liberties, but this is just to give an idea) how the neural field reduces back to the original population dynamics, which was in itself already the $N\to\infty$ limit of the network. The case of population dynamics therefore shows a trivial example of an embedding in $[0,1]$.

### Population dynamics with an infinite number of populations

When taken in isolation, the previous developments might sound a little bit silly, though they serve to give an impression of the problem at hand, and an interpretation of the rest of this work. For a finite number of populations $N_\text{pop}$, the population dynamics is perfectly defined, however what happens as $N_\text{pop} \to \infty$ ? The core of the question lies in the fact that we can easily *enumerate* a finite number of populations, but the sum $\sum_{b=1}^{N_\text{pop}}$ must have a clearly defined interpretation when this number tends to infinity.

In light of the toy model presented in @sec-nf, we can reinterpret the density $\rho$ of the neural field equation. Every point $\vec z \in \R^p$ (let us consider $p=2$ for the rest) describes an infinite population of homogeneous neurons, associated with a probability (density) $\rho(\vec z)$. Therefore, in order to write an equivalent neural field in $[0,1]$, we need a way to *enumerate* the populations. This is done using a mapping $S : [0,1]^2 \mapsto [0,1]$, where we traverse the $[0,1]^2$ space by following the order in the image $[0,1]$.

Additionally, the notion of population dynamics helps give another interpretation of the grid integration method given in @sec-gridmethod. When we discretize the $[0,1]^2$ space, we are effectively performing a coarse-graining approximation and simulating a finite number of populations. For every bin localized by $(v_1, v_2)$, we are sampling the connectivity kernel $w_U((v_1, v_2), \cdot)$ and 2D neural field $h_U((v_1, v_2), t)$ (see @fig-grid-sampling). In the rest of the chapter, we refer to these 2D bins as "square populations", and to the corresponding 1D bins as "segment populations".

![Every point of the grid effectively performs coarse-graining, and defines a point where we sample the connectivity kernel $w_U$ and the 2D neural field $h_U$](figures/photo_2023-06-10_21-13-23.jpg){#fig-grid-sampling}

For a finite number of populations $N_\text{pop}$, applying a bijective mapping $S$ will evidently give an equivalent neural field, because the connectivity matrix is permutation-invariant, and the resulting connectivity matrix $J_{S(a),S(b)}$ will therefore give equivalent dynamics. As we increase $N_\text{pop}$ by taking finer and finer grids, the image of the grids defines a sequence of mappings $S^n: \text{grid in}\ [0,1]^2 \mapsto \text{grid in}\ [0,1]$, which we aim to study in the rest of this chapter.

> How do we define a sequence of mappings $S^n$ such that the limiting $S : [0,1]^2 \mapsto [0,1]$ can express a 1D neural field with identical dynamics ?

## Defining the neural field in $[0,1]$

We start by writing the neural field equation in the $[0,1]$ space. For this, we define a measurable and bijective mapping $S : [0,1]^2 \rightarrow [0, 1]$. Furthermore, we let $\lambda : [0,1]^2 \rightarrow [0, \infty]$ be a measure on the unit square (for instance, the Lebesgue measure). Taking the composition $\lambda \circ S^{-1}$ then defines a measure on $[0,1]$, which keeps track of how 2D surfaces are transformed to 1D lengths through the mapping $S$.

The connectivity kernel $\tilde w$ defines the interactions between the populations in the $[0,1]$ embedding, which we define relating to the uniform kernel on $[0,1]^2$ as :

$$
\tilde w(\alpha, \beta) = w_U(S^{-1}(\alpha), S^{-1}(\beta))
$$ {#eq-kernel-1d}

Defining the initial condition for the 1D neural field as $\tilde h(\alpha, t=0) = h_U(S^{-1}(\alpha), t=0)$, we have fully defined the equation of evolution in the one-dimensional embedding in @eq-nf-1d.

$$
\partial_t \tilde h(t, \alpha) = -\tilde h(t, \alpha) + \int_{[0,1]} \tilde w(\alpha, \beta) \phi(\tilde h(t, \beta)) \lambda \circ S^{-1}(\d \beta)
$$ {#eq-nf-1d}

The definition of the connectivity kernel in @eq-kernel-1d might seem innocent, but the key element here is the inclusion of the inverse $S^{-1}$. We recall that in the toy model, the original connectivity kernel is written as $w(\vec z, \vec y) = \sum_{\mu=1}^p \tilde \phi(y_\mu) z_\mu$. This kernel is highly regular, by which we mean that by its continuity in each of its arguments, populations nearby in the embedding will "sense" similar connectivity functions. Since it is also differentiable, we can write this as a Taylor expansion in the first argument :

$$
\begin{aligned}
w(\vec z + \vec \epsilon, \vec y) &= w(\vec z, \vec y) + \sum_{\mu=1}^p \epsilon_\mu \frac{\partial w}{\partial z_\mu} + \bO(\norm{\vec \epsilon}^2) \\
&= w(\vec z, \vec y) + \sum_{\mu=1}^p \epsilon_\mu \phi(y_\mu) + \bO(\norm{\vec \epsilon}^2) \\
&\approx w(\vec z, \vec y) 
\end{aligned}
$$

When we do the change of variables with the CDF, we don't encounter problems because the Gaussian (inverse) CDF is differentiable and its derivative is continuous[^inversecdf]; therefore $w_U(\vec v, \vec u) = w(\textrm{CDF}^{-1}(\vec v), \textrm{CDF}^{-1}(\vec u))$ is still highly regular.

[^inversecdf]: Strictly speaking, we have the problem that $\textrm{CDF}^{-1}(v) \xrightarrow{v \to 1} \infty$, and so the derivative diverges when $v \to 1$ (and similarly, when $v \to 0$). However, this is not a problem, since these limits correspond to $z \to \pm \infty$, where the density vanishes.

Therefore, we can see why the composition of $w_U$ with a possibly non-differentiable mapping $S$ might be problematic. Since the only thing we know about $S$ is that it is measurable and bijective, *there is no guarantee that the regularity of the kernel is conserved through the composition with the mapping $S$.*

## Mappings of $[0,1]^2$ to $[0,1]$

### The search for differentiable mappings

We start our search of mappings by attempting to find a mapping $S$ that is differentiable.
The easiest example of such a mapping would be a linear projection, for instance, $P(v_1, v_2) = v_1$. This mapping is obviously continuous and differentiable in both components, and surjectively maps the unit square to the unit segment.

However, the obvious problem with this approach is that linear projections are not invertible. Hence, should we nevertheless define an inverse $P^{-1}$ that computes the preimage of $\alpha \in [0,1]$, we have that

$$
P^{-1}(\alpha) = \{(\alpha, v_2)\ |\ v_2 \in [0,1] \}.
$$

In other words, $P^{-1}(\alpha)$ defines a vertical line in the $[0,1]^2$ embedding. When we then take the composition $w_U \circ P^{-1}$, the regularity is destroyed, in the sense that the same population $\alpha$ in the 1D embedding will "sense" all the connectivity kernels on the vertical line.

A more sophisticated attempt, in the spirit of @sec-mapping-cdf, might be to consider the joint cumulative function

$$
\alpha = \mathrm{CDF}(z_1, z_2) = \int_{-\infty}^{z_1} \int_{-\infty}^{z_2} \rho(y_1, y_2) \d y_1 \d y_2.
$$

However, this again fails because the joint cumulative function is not bijective, and isolines of the CDF define diagonals in the uniform embedding space $[0,1]^2$.[^jointcdf]

[^jointcdf]: We note to avoid any confusion, that in @sec-mapping-cdf the CDF is taken component-wise, instead of jointly. The univariate Gaussian CDF defines a bijection between $\R$ and $(0,1)$, but the joint CDF does not.

Perhaps the linear projection or CDF examples are a bit too naive, and we need to look for more complex examples. We might ask the question of whether a diffeomorphism (a differentiable bijection with a differentiable inverse) even exists between the unit square and the unit segment. Unfortunately, results from topology give a negative answer, and even worse, it can be shown that there is no homeomorphism (a continuous bijection with continuous inverse) between the unit square and the unit segment. This is stated by Netto's theorem, which imposes continuous bijections to conserve the dimensionality between the domain and the image.

### Mappings of $[0,1]$ into $[0,1]^2$

Since in the composition $w_U \circ S^{-1}$ we work with the inverse mapping, it might be judicious to directly consider functions from $[0,1]$ into $[0,1]^2$. We have already established that there is no homeomorphism between those spaces, and need to relax one of the conditions.

Functions of $[0,1]$ into $[0,1]^2$ are extensively discussed in chapter 1 of @kharazishvili_strange_2017, and we cite two classes of such functions :

1. Cantor functions are bijections between $[0,1]$ and $[0,1]^2$. By the non-existence theorem, these functions cannot be continuous.
2. Peano functions are continuous surjections from $[0,1]$ to $[0,1]^2$. They cannot be injections. For reasons that will become apparent in the following sections, Peano functions are often called space-filling curves.

In light of these results, we might be tempted to search for expressions for Cantor-type functions, since having a bijection would allow us to have a direct mapping between the square and the segment populations. Unfortunately, these functions are discontinuous which breaks the regularity of the kernel.

Peano functions seem like good candidates for the mapping $S^{-1}$, and are extensively studied in the literature (we refer for instance to [@sagan_space-filling_1994;@Sagan1993]). They have the nice property that they are continuous and surjective, so we the segment populations can "reach" the entirety of the square populations. All things considered, non-injectivity is not a problem, since it just means that two segment populations can "refer" to the same square population, which does not break the regularity of the kernel, since the two segment populations can just have their own neighborhoods. Furthermore, we will see that in our examples, the non-injectivity can be ignored because the mapping is quasi-bijective.

![Non-injectivity of $S^{-1}$ means that two segment populations can refer to the same square population.](figures/photo_2023-06-12_00-46-26.jpg)

### Sequences of bijective mappings

#### Mappings as binary expansions {#sec-mappings-binary}

We consider $(v_1, v_2) \in [0,1]^2$. Given $n \in \mathbb N$, we write a (truncated) binary expansion of $v_1$ and $v_2$.

$$
\begin{aligned}
v^{(n)}_1 &= \sum_{l=1}^{n} b^1_l 2^{-l} = 0.b^1_1b^1_2\cdots b^1_{n} \iff b^1_l = \mathrm{Ind}\left\{2^{l-1}v_1 - \lfloor2^{l-1}v_1\rfloor \geq \frac{1}{2}\right\} \\
v^{(n)}_2 &= \sum_{l=1}^{n} b^2_l 2^{-l} = 0.b^2_1b^2_2\cdots b^2_{n} \iff b^2_l = \mathrm{Ind}\left\{2^{l-1}v_2 - \lfloor2^{l-1}v_2\rfloor \geq \frac{1}{2}\right\}
\end{aligned}
$$ {#eq-trunc}

The intuition behind this expansion is we take a segment, split it in two, write 0 to go left, write one to go right, then recurse. Then combine the bits of $v^{(n)}_1$ and $v^{(n)}_2$ to get the position in the $[0,1]^2$ embedding.

![Splitting $[0,1]$ recursively gives the binary bit expansion of $v_1$](figures/photo_2023-06-10_17-21-28.jpg)

We define the mapping $S^n$ that takes the $n$-bit truncations $v^{(n)}_1$ and $v^{(n)}_2$ as input, and outputs a $2n$ bit truncation $\alpha^{(n)}$ of $\alpha=S(v_1, v_2)$, where $S$ is the $n\to\infty$ limit of the sequence $S^n$.

$$
\begin{alignat}{2}
S^n : &\{0, \tfrac{1}{2^n}, \cdots, 1-\tfrac{1}{2^n}\}^2 &&\mapsto \{0, \tfrac{1}{4^n}, \cdots, 1-\tfrac{1}{4^n}\} \\
&(v^{(n)}_1,v^{(n)}_2)=(0.b^1_1 b^1_2 \cdots b^1_n, 0.b^2_1 b^2_2 \cdots b^2_n) &&\mapsto \alpha^{(n)}=0.b_1 b_2 \cdots b_{2n}
\end{alignat}
$$ {#eq-Sn}

In general, each bit $b^k$ of the image would be a binary function of the $n$ bits of $v_1$ and the $n$ bits of $v_2$, implicitly $b_k = b_k(b^1_1, b^1_2, \cdots, b^1_n, b^2_1, b^2_2, \cdots, b^2_n)$.
With this, we can formulate any mapping of $4^n$ distinct elements (here : squares of size $2^{-n} \times 2^{-n}$) onto $4^n$ elements (here : segments of length $4^{-n}$).

The intuition behind the use of $2n$ bits for the image is that we take $2^n \times 2^n = 4^n$ squares in $[0,1]^2$ and map those to $4^n$ segments in $[0,1]$. Informally, we map $n+n$ bits onto $2n$ bits.

In the following, we will restrict ourselves to mappings where the functions $b_k(\{b^1_l\}_{l\in\{1,\cdots,n\}}, \{b^2_l\}_{l\in\{1,\cdots,n\}})$ are defined in such a way that each bit of the inputs is used once, and only once. Therefore, for each bit $k$ there is a function $b_k(b^{\mu_k}_{l_k})$ acting on only one bit of the input. $\mu_k \in \{1,2\}$ is the input component and $l_k \in \{1,\cdots,n\}$ the bit number corresponding to $k$. Since this function is binary, the only allowed definitions are

* $b_k(b^{\mu_k}_{l_k}) = b^{\mu_k}_{l_k}$ (identity),
* $b_k(b^{\mu_k}_{l_k}) = 1-b^{\mu_k}_{l_k}$ (inversion),
* $b_k(b^{\mu_k}_{l_k}) = 0$ (constant zero),
* $b_k(b^{\mu_k}_{l_k}) = 1$ (constant one).

The constant functions are ignored, since they effectively just restrict the output domain to a subset of $[0,1]$; and since only $2n$ bits are allowed for the output, they are a "waste" by ignoring information of the input.
Because they don't add information to the output (they just "swap" the values of the output bits at known locations), inversion functions are also not considered.

All this mathematical formalism expresses, in essence, that the mapping $S^n$ is just a *reordering* of the input bits. $S^n$ takes $4^n$ points embedded in $[0,1]^2$, and reorders them onto $4^n$ points in $[0,1]$.

Therefore, there is a one-to-one mapping between the points on the square and the points on the segment, and the mapping $S^n$ is bijective.

These simplifications might seem too restrictive, however, we will show in the rest of this thesis that this formulation is flexible enough to express mappings that have good properties such as "locality", all while being simple enough to help build intuitions.

TODO : might need to talk about how $S(v^{(n)}_1,v^{(n)}_2) = S^n(v^{(n)}_1,v^{(n)}_2)$ in appendix ? so basically the limit $n\to\infty$ is the limit where we take infinite precision in the input. Or we can drop the $n$ on the inputs, $\alpha^{(n)} = S^n(v^{(n)}_1,v^{(n)}_2) = S^n(v_1,v_2)$ (i think this true just for pointwise convergent mappings)

TODO NOTE : talk about how we center the bins instead of taking the corners. this is not super important in the n -> oo limit, but numerically at 0 and 1 the CDF-1 goes to infty. at infinity, the connectivity kernel is infinite but we have zero neurons (zero measure) (linear growth vs exponential tail of the gaussina) : effectively the recurrent current is zero and we can safely ignore these effects.

#### A link with space-filling curves

With this formalism in mind, we give a geometric interpretation to the mappings $S^n$. The core idea is that the populations in $[0,1]$ have a clear ordering to them, and we can enumerate them simply by their position in the embedding. Since we constructed $S^n$ to be bijective, this enumeration traces a path in the corresponding populations of the $[0,1]^2$ embedding.

We illustrate this by introducing the "column" mapping.

$$
\alpha^{(n)} = C(v^{(n)}_1,v^{(n)}_2) = 0.b^1_1 b^1_2 \cdots b^1_n b^2_1 b^2_2 \cdots b^2_n
$$ {#eq-column}

The intuition behind this mapping is we enumerate the populations on a grid column by column, letting the index of the rows vary the fastest. We visualize this mapping in @fig-col, by coloring the square populations by the position of the corresponding $[0,1]$ population, and additionally drawing lines between the points to help guide the eye.

```{python}
#| label: fig-col
#| fig-cap: Column mapping for $n=2$. Numbers represent the ordering of the populations in the $[0,1]$ embedding, neighbouring populations are connected by lines.
#| echo: false
def plot_trace(mapping_cls, nrec, do_trace: bool = True, do_numbers: bool = True):
	F = ColumnMapping.new_nrec(nrec).inverse_samples(centered=True)
	mapping = mapping_cls.new_nrec(nrec)
	alpha = mapping(F)

	fig, ax = plt.subplots(facecolor='#FFFFFF00')
	im = ax.imshow(alpha.reshape((2**nrec, 2**nrec)).T, origin='lower', vmin=0, vmax=1, extent=(0,1,0,1), interpolation='none')
	mapping_letter = { ZMapping: 'Z', ColumnMapping: 'C', AntiZMapping: 'A' }.get(mapping_cls, 'S')
	plt.colorbar(im, ax=ax, label=f'$\\alpha={mapping_letter}^{{n}}(v_1, v_2)$')
	if do_trace:
		# plot the trace
		ax.plot(*F[np.argsort(alpha)].T)
		# hide trace at numbering location
		ax.scatter(*F.T, s=250, c=alpha, cmap='viridis', vmin=0, vmax=1, zorder=10)  
	# add the bin numbers
	if do_numbers:
		for xy, i in zip(F[np.argsort(alpha)], range(1, 4**nrec+1)):
			ax.annotate(i, xy=xy, va='center_baseline', ha='center', zorder=20, color='white', fontweight='bold',
				path_effects=[pe.withStroke(linewidth=1, foreground='black')])
	ax.set_aspect('equal')
	ax.set_xlabel('$v_1$'); ax.set_ylabel('$v_2$')
	ax.set_xlim((0,1)); ax.set_ylim((0,1))
	ax.xaxis.set_major_locator(ticker.MultipleLocator(1/2**2))
	ax.yaxis.set_major_locator(ticker.MultipleLocator(1/2**2))
	# for x in np.linspace(0, 1, 2**nrec+1): ax.axvline(x, ymin=0, ymax=1, color='tab:gray', linewidth=0.5, zorder=0, clip_on=False)
	# for y in np.linspace(0, 1, 2**nrec+1): ax.axhline(y, xmin=0, xmax=1, color='tab:gray', linewidth=0.5, zorder=0, clip_on=False)
	breathe()
	return fig, ax

fig, ax = plot_trace(ColumnMapping, nrec=2)
display(fig)
```

For finite $n$, we write the inverse of the column mapping in @eq-column-inverse.

$$
\begin{aligned}
v^{(n)}_1 = C^{-1}_1(\alpha^{(n)}) &= 0.b_1 b_2 \cdots b_n = \sum_{k=1}^n b_{k} 2^{-k} \\
v^{(n)}_2 = C^{-1}_2(\alpha^{(n)}) &= 0.b_{n+1} b_{n+2} \cdots b_{2n} = \sum_{k=1}^n b_{n+k} 2^{-k}
\end{aligned}
$$ {#eq-column-inverse}

This way of enumerating the square populations might be familiar to people who have worked with numerical arrays. In such (contiguous) arrays, the values are laid out in memory in "row-major" order (C-style) or "column-major" order (Fortran-style); and we can go from the 1D memory representation to a 2D matrix representation by using "reshape" operations. Effectively, this is how this mapping is implemented in code, using `numpy.ravel_multi_index` (2D to 1D) and `numpy.unravel_index` (1D to 2D).

#### "Local" space-filling curves {#sec-local}

Research in computer science and efficient data structures has given rise to other ways of storing 2-dimensional information in a 1-dimensional memory.
A problem in GPU computing is that of texture locality: sampling textures often involves reading data corresponding to a small 2D region, and therefore to improve speed (minimize cache misses), the 2D data should be packed close together in the 1D memory.
One way of ordering the 2D texture data in the 1D memory is by using the Z-order curve (also known as "Morton mapping" or "Lebesgue curve" @Sagan1994). In this way, points close in 2D *tend* to be tightly packed in 1D. We can define the Z-order curve (Z-mapping) in the following way[^zcurve] :

[^zcurve]: The Z-curve is often defined in a transpose manner, putting first the bits of $v_2$ (vertical component) then the bits of $v_1$, instead of $v_1$ then $v_2$. By doing this, we get the "upright" Z shapes, as [shown here](https://en.wikipedia.org/wiki/Z-order_curve#/media/File:Four-level_Z.svg).

$$
\alpha^{(n)} = Z(v^{(n)}_1,v^{(n)}_2) = 0.b^1_1 b^2_1 b^1_2 b^2_2 \cdots b^1_n b^2_n = \sum_{k=1}^{n} b^1_k 2^{1-2k} + b^2_k 2^{-2k}
$$ {#eq-z}

And its inverse is :

$$
\begin{aligned}
v^{(n)}_1 = Z^{-1}_1(\alpha^{(n)}) &= 0.b_1 b_3 \cdots b_{2n-1} = \sum_{k=1}^n b_{2k-1} 2^{-k} \\
v^{(n)}_2 = Z^{-1}_2(\alpha^{(n)}) &= 0.b_2 b_4 \cdots b_{2n} = \sum_{k=1}^n b_{2k} 2^{-k}
\end{aligned}
$$ {#eq-z-inverse}

@fig-z is a visualization for the Z-mapping when $n=2$, and shows how it is composed of the "Z" shapes giving it its name.

```{python}
#| label: fig-z
#| fig-cap: Z-mapping for $n=2$. Numbers represent the ordering of the populations in the $[0,1]$ embedding, neighbouring populations are connected by lines.
#| echo: false
fig, ax = plot_trace(ZMapping, nrec=2)
display(fig)
```

To illustrate this notion of locality, we take a few consecutive populations in 1D and show where the corresponding populations fall in the 2D embedding. This is done in @fig-locality, in which we see that the Z-mapping seems to conserve locality from 1D to 2D: populations close in 1D seem to be close in 2D.

We also see that the Column mapping is not local in this sense, since there is a large variation along the vertical direction, which seems to always be "of order one". This can be explained by looking at the expression for $C^{-1}_2(\alpha^{(n)}) = 0.b_{n+1} b_{n+2} \cdots b_{2n}$ : small variations of the order of $2^{-n}$ in the value of $\alpha$ always result in variations of order one in $C^{-1}_2$.

```{python}
#| label: fig-locality
#| fig-cap: Locality of the mappings. The 2D populations corresponding to three small segments of 1D populations of length $\tfrac{1}{16}$ are shown. The figures are generated with $n=6$.
#| echo: false
##| fig-subcap:
##|   - Z mapping
##|   - Column mapping
##| layout-ncol: 2
fig, axes = plt.subplot_mosaic('aa\nbc', facecolor='#FFFFFF00', figsize=(6, 4), height_ratios=[1, 5], subplot_kw=dict(facecolor='#FFFFFF00'))
nrec = 6

cmap = mpl.colormaps['viridis']
cmap.set_bad(color='#FFFFFF00')

alpha = np.linspace(0, 1, 4**nrec+1)
mask1d = (0.2 <= alpha) & (alpha < 0.2+1/4**2)
mask1d |= (0.8 <= alpha) & (alpha < 0.8+1/4**2)
mask1d |= (0.5 <= alpha) & (alpha < 0.5+1/4**2)
alpha[~mask1d] = np.nan
axes['a'].imshow(alpha[None, :], vmin=0, vmax=1, extent=((0,1,0,0.2)), interpolation='none', cmap=cmap)
axes['a'].set_aspect('auto')
axes['a'].spines.left.set_visible(False)
axes['a'].yaxis.set_visible(False)
axes['a'].set_title('Populations in 1D', fontsize='medium')
axes['a'].set_xlabel('$\\alpha$')
breathe(axes['a'])

for mapping_cls, ax in zip((ZMapping, ColumnMapping), (axes['b'], axes['c'])):
	F = ColumnMapping.new_nrec(nrec).inverse_samples(centered=True)
	mapping = mapping_cls.new_nrec(nrec)
	alpha = mapping(F)
	mask1d = (0.2 <= alpha) & (alpha < 0.2+1/4**2)
	mask1d |= (0.8 <= alpha) & (alpha < 0.8+1/4**2)
	mask1d |= (0.5 <= alpha) & (alpha < 0.5+1/4**2)
	alpha[~mask1d] = np.nan
	ax.imshow(alpha.reshape((2**nrec, 2**nrec)).T, origin='lower', vmin=0, vmax=1, extent=(0,1,0,1), interpolation='none', cmap=cmap)
	ax.set_xlabel('$v_1=S_1^{{-1}}(\\alpha)$'); ax.set_ylabel('$v_2=S_2^{{-1}}(\\alpha)$')
	ax.set_xlim((0,1)); ax.set_ylim((0,1))
	ax.xaxis.set_major_locator(ticker.MultipleLocator(1/2**2))
	ax.yaxis.set_major_locator(ticker.MultipleLocator(1/2**2))
	ax.set_title(f'To 2D with {mapping_cls.__name__}', fontsize='medium')
	breathe(ax)

display(fig)
```

The attentive reader might however object to this treatment of locality, since in some sense the Column mapping is local, but from 2D to 1D. This can be seen from the definition of $\alpha^{(n)}=C(v^{(n)}_1,v^{(n)}_2) = 0.b^1_1 b^1_2 \cdots b^1_n b^2_1 b^2_2 \cdots b^2_n$, since small variations of order $2^{-n}$ in $v^{(n)}_1$ result in small variations also of order $2^{-n}$ in $\alpha^{(n)}$. Similarly, small variations of order $2^{-n}$ in $v^{(n)}_1$ result in even smaller variations also of order $2^{-2n}$ in $\alpha^{(n)}$ ! We therefore seem to have two notions of locality: from 1D to 2D, and from 2D to 1D; and the Column mapping shows that they are not equivalent.

However, we argue that in our case only the notion of locality from 1D to 2D matters. In essence, the dynamics in the 2D embedding are already known, and we ask the question of whether it is possible to write an equivalent neural field 1D (in the sense that it has identical latent trajectories). Without repeating the intuition given in @sec-intro-simnf, if neighboring populations in 1D have similar potentials, then this allows us to write a neural field equation in $[0,1]$. Therefore we would like to have the property that populations close in 1D are (inversely) mapped to populations close in 2D, which already have similar potentials.

<!-- note for me : find a formulation for the recursive local mapping. it is not trivial, since it has

```
n=1
0 0 -> 00
1 0 -> 01
1 1 -> 10
0 1 -> 11

n=2
00 00 -> 0000
01 00 -> 0001
01 01 -> 0010
00 01 -> 0011
10 00 -> 0100
11 00 -> 0101
11 01 -> 0110
10 01 -> 0111
10 10 -> 1000
11 10 -> 1001
11 11 -> 1010
10 11 -> 1011
00 10 -> 1100
01 10 -> 1101
01 11 -> 1110
00 11 -> 1111
```

this might be nice as an appendix, to show that our formulation is very simplistic, and more complex formulations can also have nice properties

this also shows our proof is "incomplete", because it relies on the fact we have a permutation of the bits, although it is not strictly necessary. we have a sufficient, but not necessary proof

also : this might be a type of peano curve or smth. see @Sagan1994

https://www.cut-the-knot.org/Curriculum/Geometry/LebesgueCurve.shtml
-->

### Sequence of mappings and their limit

#### Pointwise convergence of mappings

We now discuss what happens to the bijective mappings when $n\to\infty$. For this, let us consider again the "random mapping" considered in the introduction @sec-intro-simnf. We discussed how this hypothetical mapping maps each position in $[0,1]^2$ into a "random position" in $[0,1]$. A finite-$n$ formulation would be that the mapping corresponds to a *random permutation* of the bits. It is unclear how to formally write this mapping in a way that is well-defined for a given $n$, since by definition, it is random. For a given position of the input $(v_1,v_2)$, we would like that the image is stable in the $n\to\infty$ limit.

$$
S^{n}(v_1, v_2) \xrightarrow{n\to\infty} \alpha =: S(v_1, v_2)
$$

This is the condition of "pointwise convergence" of the mappings. We might be tempted to write that any mapping for which we can write the binary expansion $S^n$ is pointwise convergent, however, we show by the example of the "anti-Z" mapping that this is not the case. Let us define for finite $n$

$$
\begin{aligned}
\alpha^{(n)} &= A(v^{(n)}_1,v^{(n)}_2) = 0.b^1_n b^2_n b^1_{n-1} b^2_{n-1} \cdots b^1_1 b^2_1 = \sum_{k=1}^{n} b^1_{n+1-k} 2^{1-2k} + b^2_{n+1-k} 2^{-2k} \\
\iff& \\
v^{(n)}_1 &= A^{-1}_1(\alpha^{(n)}) = 0.b_{2n-1} b_{2n-3} \cdots b_{1} = \sum_{k=1}^n b_{2(n-k)-1} 2^{-k} \\
v^{(n)}_2 &= A^{-1}_2(\alpha^{(n)}) = 0.b_{2n} b_{2n-2} \cdots b_{2} = \sum_{k=1}^n b_{2(n+1-k)} 2^{-k}
\end{aligned}
$$

```{python}
#| label: fig-anti-z
#| fig-cap: Anti-Z mapping for $n=2$. Numbers represent the ordering of the populations in the $[0,1]$ embedding. We don't plot the trace, because it has many crossovers and overlaps which makes the plot hard to read.
#| echo: false
fig, ax = plot_trace(AntiZMapping, nrec=2, do_trace=False)
display(fig)
```

@fig-anti-z gives some insight into the problem of this mapping. For a fixed input $(v_1,v_2)$, we see that the output point position jumps around and that these jumps do not decay, they are always of order 1. The intuition behind pointwise convergence is that as $n$ increases, the image of a point by $S^n$ is continually refined.

Another way of looking at this is to consider how $\alpha^{(n)}$ fluctuates as the precision of the input increases with $n \to \infty$. As we refine the input with fluctuations of order $2^{-n}$, the output $\alpha^{(n)} = b^1_n 4^{-1} + b^2_n 4^{-2} + b^1_{n-1} 4^{-3} + b^2_{n-1} 4^{-4} \cdots$ fluctuates (at least) with amplitude $\sim 4^{-2}$ independantly of $n$. Therefore the output never converges, and the mapping $A^{n}$ is not pointwise convergent.

In our setting, how can we guarantee that the mappings are pointwise convergent? We need to show that small fluctuations in the input result in small fluctuations in the output. The binary expansion allows us to argue that for the sequence of mappings to be pointwise convergent, we need that the "least significant bits" (LSB, the bits to the right of the binary expansion) of the input are mapped to the LSB of the output. Conversely, the "most significant bits" (MSB, the bits to the left of the binary expansion) should be mapped to the MSB of the output.

We illustrate this with the Z-mapping, written in @eq-z. Corrections of order $2^{-n}$ in $v_1$ and $v_2$ induce corrections of order $2^{1-2n}$ and $2^{-2n}$ in $\alpha$ respectively. Therefore, it is easy to see that the Z-mapping is pointwise convergent as $n\to\infty$. Similarly, the Column mapping in @eq-column is also pointwise convergent, since as discussed @sec-local it has a locality of 2D to 1D.

#### Conservation of bijectivity in the $n\to\infty$ limit {#sec-ninfty-bijection}

However, the behavior of the Column mapping and that of the Z-mapping are very different in the $n\to\infty$ limit. Up until now, we have avoided this issue by considering the finite-$n$ approximation of the mappings, which by construction is bijective. We illustrate the infinite-$n$ (numerically, large $n$) limit of these two mappings in @fig-n-to-infty.

```{python}
#| label: fig-n-to-infty
#| fig-cap: The $n\to\infty$ limit of mappings.
#| fig-subcap:
#|   - Z-mapping
#|   - Column mapping
#| layout-ncol: 2
#| echo: false
fig, ax = plot_trace(ZMapping, nrec=6, do_trace=False, do_numbers=False)
display(fig)
fig, ax = plot_trace(ColumnMapping, nrec=6, do_trace=False, do_numbers=False)
display(fig)
```

We see that the Column mapping converges to projection on $v_1$, in other words, $\alpha=C(v_1, v_2) = v_1$, showing that despite every $C^{n}$ being bijective, the limit of the sequence is not bijective. This becomes clear when looking at the binary expansion of $C$ :

$$
\begin{aligned}
C(v_1, v_2) &= \lim_{n\to\infty} C^{(n)}(v_1,v_2)\\
&= \lim_{n\to\infty} 0.b^1_1 b^1_2 \cdots b^1_n b^2_1 b^2_2 \cdots b^2_n \\
&= \lim_{n\to\infty} 0.b^1_1 b^1_2 \cdots \\
&= v_1
\end{aligned}
$$

The limit is not bijective, because the bits of $v_2$ are "lost" in the $n\to\infty$ limit, and the inverse $C_2^{-1}(\alpha)$ is not well-defined, since it "takes bits from $\infty$ to $2\infty$".

The Z-mapping however seems to converge to a fractal curve. Numerically, there seems to be an infinite number of discontinuities, but the size of these discontinuities is overwhelmingly small, and *on average* the size of these numerical discontinuities vanishes as $n\to\infty$. We won't go further into detail about this in this section (in @sec-zcurve-lebesgue we develop the convergence to a space-filling fractal curve), but this gives some intuition for the arguments presented in @sec-locality.

We finish this discussion on the limit of the sequence of mappings by asking if there are other mappings similar to the Z-mapping, in the sense that they are pointwise convergent and the limit is bijective.[^zmapping-notbijective] In an informal way, we established two conditions for this in the context of the mappings described in @sec-mappings-binary :

[^zmapping-notbijective]: Despite what can often be found in online resources, the Z-mapping is *not* a bijection because that would violate Netto's theorem. The reason for this is the ambiguity in the writing of binary expansions, and in @sec-zcurve-noninjective we give proof that the Z-mapping is not injective in the $n\to\infty$ limit. However, the Z-mapping is *quasi-bijective*, in the sense that the set of points for which the Z-mapping is not injective has zero measure, and can therefore be ignored.

1. condition of pointwise convergence: the LSB of the input must be mapped to the LSB of the output,
2. condition of "no information loss": the MSB of the input must be mapped to the MSB of the output.

These two conditions do not leave a lot of flexibility in the design of new mappings. We can evidently define mappings where we interleave the bits in a different order, such as $0.b^2_1 b^1_1 b^2_2 b^1_2 \cdots$; or mappings where we reorder the $k$ bits of the input, as in $0.b^1_2 b^2_2 b^1_1 b^2_1 b^1_4 b^2_4 b^1_3 b^2_3 \cdots$, but the conditions restrict the reordering to be permutations of neighboring bits. In essence, these mappings are not fundamentally different, and the Z-mapping can be seen as a "canonical representation" of the pointwise convergent mappings with bijective limits.

### Visualizating the mappings of the dynamic 2D neural field

It is time for a little treat. Before we argue about how to simulate the dynamics of the 1D neural field, let us simulate the dynamics of the 2D neural field, then at each timestep, we map the 2D neural field to the 1D embedding. This will give us a reference of how the 1D neural field is supposed to look under the condition that its latent dynamics are identical.

![(TODO) Animation showing the image of the dynamics 2D neural field through different mappings $S : [0,1]^2 \mapsto [0,1]$](figures/photo_2023-06-10_17-21-19.jpg)

## Coarse-graining the neural field in $[0,1]$

### Numerical aspects of implementing the mappings

In the previous sections, we have discussed what mappings have good properties, and have given a few hints as to how we can implement them numerically. Since the mappings we presented can be described as a binary expansion, they naturally lend themselves to numerical approximations. Furthermore, this is made especially easy since the mappings we defined simply are permutations of bits (in the $n < \infty$ case), and so we can write a one-to-one mapping between (unsigned integers).

Without going into too much detail on the programming, we have implemented the mappings as all inheriting from the same abstract class, and each mapping simply has to define a method `indices_to_indices2d(indices: array[int, 1]) -> array[int, 2]` to go from 1D to 2D ($S^{-1}$), and a method `indices2d_to_indices(indices2d: array[int, 2]) -> array[int, 1]` to go from 2D to 1D.[^indices] The abstract class simply defines additional quality-of-life functionality, which makes use of these two methods.

[^indices]: In the source code, only the method `indices(v12: array[float, 2]) -> array[int, 1]` is defined. This is because some mappings are implemented (for instance, the `RecursiveLocalMapping`, that we don't discuss here), which don't lend themselves to the simple permutation of bits formulation.

### Numerical approximations of a fractal structure

In the continuation of the simulation methods established in the @sec-simnf, we adopt a grid discretization scheme, as it naturally lends itself to be applied with the mappings. We however note that using binning methods, the Monte-Carlo simulation can still be applied. The intuition behind this is that sampling populations inside a small square of size $2^{-n} \times 2^{-n}$ gives rise to the notion of an "average connectivity kernel", and in our case the notion of "mean patterns". More details are given in @sec-mean-patterns.

We now address the elephant in the room: how do we numerically simulate a neural field defined on -- what appears to be -- a fractal curve? Naively, we could just proceed as follows :

1. Create a grid of $2^n \times 2^n = 4^n$ squares of size $2^{-n} \times 2^{-n}$ in the $[0,1]\times[0,1]$ embedding. Each square is located by a position $\vec{v_i} = (v_{1,i},v_{2,i}),\,i\in\{1,\cdots\,2^n\}$.
2. Compute a discretized version of the 1D connectivity kernel
   $$
   \begin{aligned}
   \tilde w(\alpha_i, \beta_j) &= \tilde J_{\alpha_i,\beta_j}\\
   &= \tilde J_{S(i),S(j)}\\
   &\stackrel{\text{def}}{=} J_{ij}\\
   &= w_U(\vec{v_i}, \vec{v_j})
   \end{aligned}
   $$
   at every square location (where we denote for simplicity of notation $S(i)=S(\vec{v_i})$).
3. Simulate the neural field in $[0,1]$ on the corresponding discretized space with discretization step $4^{-n}$.

The problem with this approach is that since the mappings are just *permutations* of the square populations, the resulting connectivity matrix $\tilde J_{\alpha_i,\beta_j}$ is just a permutation of the rows and columns of $J_{ij}$, and will therefore trivially lead to the identical neural field dynamics.

In some sense, we would like to "take into account" the fact that a mapping of the square populations has been performed. Let us consider what happens to the mapping when $n$ is large, and in particular how $S^{n/2}$ and $S^{n}$ behave. Since the mappings are pointwise convergent and we are considering large $n$, we argue that both $S^{n/2}$ and $S^{n}$ are good approximations of the true mapping $S$. More importantly, $S^{n/2}$ can be considered as an approximation of $S^{n}$. This is interesting from a numerical standpoint since it is feasible to compute $S^{n}$, and the approximation $S^{n/2}$ can be obtained by "coarse-graining" $S^n$. Another way of thinking about this is that $S^n$ is considered to be the "mathematically exact" fractal structure and $S^{n/2}$ the numerical approximation.

The way we go from $S^n$ to $S^{n/2}$ is by averaging the segment populations by batches of $2^n$, as schematized in @fig-averaging. We call this operation "downsampling" the $S^n$ mapping.

![(TODO) Averaging $2^n$ of the $4^n$ segments populations of $S^n$ results in the $2^n = 4^{n/2}$ segments populations of $S^{n/2}$](figures/photo_2023-06-13_12-15-32.jpg){width=400px #fig-averaging}

The geometrical intuition we can give for this is how the averaging of the segment populations leads to sampling the connectivity kernel at the "average" positions in the $[0,1]^2$ embedding. In @fig-averaging-geom, we color-code the square populations corresponding to the segment populations before averaging.

::: {#fig-averaging-geom layout-ncol=2}

![Z-mapping](figures/averaging_z.png)

![Column mapping](figures/averaging_column.png)

(TODO) The corresponding square populations corresponding to the segment populations before downsampling.
:::

One question that may arise is why we take $n/2$ instead of $n-1$. The reason for this is that we want the number of averaged segment populations to scale with $n$. If we had chosen $S^{n-1}$, then the number of segment populations is related via

$$
N_\text{pop}(n-1) = 4^{n-1} = 4^{n} / 4 = N_\text{pop}(n) / 4.
$$

In other words, we would be averaging a constant number (4 in this case) of segment populations, independently of $n$. Choosing a fraction of $n$ for the approximation then allows us to ensure that we average $2^n$ populations, since

$$
N_\text{pop}(n/2) = 4^{n/2} = 4^n / 4^{n/2} = 4^{n} / 2^{n} = N_\text{pop}(n) / 2^n.
$$

Furthermore, this number of averaged populations still vanishes, since

$$
N_\text{pop}(n/2) / N_\text{pop}(n) = 2^{-n} \xrightarrow{n\to\infty} 0.
$$

## Simulations of neural fields in 3D, 2D and 1D

### Application of $S$ : from $[0,1]^2$ to $[0,1]$

![](figures/coarsegraining.png)

### Iterative application of $S$ : from $[0,1]^3$ to $[0,1]$

#### Continuity on one axis and fractal on the other: from $[0,1]^3$ to $[0,1]^2$

#### From fractal to fractal : $[0,1]^2$ to $[0,1]$

![Effective sampling points after averaging](figures/mapping_3d_2d_1d.png)

![Equivalence of dynamics, from 3D to 2D to 1D](figures/embedding_3d_2d_1d.mp4)