### 2023-06-15 manuscript/index.qmd

::: {.content-visible when-format="html"}
::: {.content-hidden}
# Home {.unnumbered}
:::

![Smooth neural fields in $p$ dimensions can be mapped to equivalent neural fields in one dimension, and the dynamics are equivalent.](chapters/figures/photo_2023-06-15_13-51-39.jpg)
:::