### 2023-06-16 manuscript/appendices/B_zcurve.qmd
dkfjwen fhjk !!!!!

::: {.content-hidden}
{{< include ../macros.tex >}}
\def\C{\mathcal{C}}
\def\P{\mathcal{P}}
\def\B{\mathcal{B}}
:::

# The Z-curve is not injective {#sec-zcurve-noninjective}

Let us study, in a bit more detail, the properties of the Z-curve $Z^{-1}$ introduced in @sec-local, and, by extension, the properties of the Z-mapping $Z$.

## Ambiguity of binary expansions

Although it might seem that the Z-curve defines a bijection (this is often falsely relayed in online resources, and was even initially believed by Cantor when he discovered this function, before being corrected by Dedekind @fleron1994cantorhistory), this is not the case.

The problem with the definition is the ambiguity of (binary) expansions. If we allow infinite sum expansions, then the two representations of $\tfrac 12$ are equivalent:

$$
\begin{aligned}
\frac 12 &= 0.1000 \cdots = \sum_{l=1}^\infty b_l 2^{-l},\ \text{where}\ b_1=1, b_{l>1}=0 \\
&= 0.0111 \cdots = \sum_{l=1}^\infty b^\prime_l 2^{-l},\ \text{where}\ b^\prime_1=0, b^\prime_{l>1}=1
\end{aligned}
$$

Although $Z^n$ are still bijections, this ambiguity is problematic for the $n\to\infty$ limit, since we can find multiple pre-images for the same $\alpha$ mapped by $Z$. This illustrate this, let us take a finite integer $m$, and write a counter-example of bijectivity. Similarly to the main text, we define the binary expansions of $v_1$ and $v_2$:

$$
\begin{aligned}
v_1 &= \sum_{l=1}^\infty b^1_l 2^{-l} = 0.b^1_1b^1_2 \cdots b^1_{m} b^1_{m+1} \cdots \\
v_2 &= \sum_{l=1}^\infty b^2_l 2^{-l} = 0.b^2_1b^2_2 \cdots b^2_{m} b^2_{m+1} \cdots
\end{aligned}
$$

We now consider the $l > m$ binary digits of $v_1$ to give ambiguous representations of $v_1^{(m)} + 2^{-m-1}$, where $v_1^{(m)}$ is the finite-$m$ truncation of $v_1$.

$$
\begin{aligned}
v_1 &= 0.b^1_1b^1_2 \cdots b^1_{m} 1000 \cdots \\&\qquad\implies \alpha = 0.b^1_1 b^2_1 \cdots b^1_m b^2_m 1 b^2_{m+1} 0 b^2_{m+2} 0 b^2_{m+3} \cdots \\
&= 0.b^1_1b^1_2 \cdots b^1_{m} 0111 \cdots \\&\qquad\implies \alpha^\prime = 0.b^1_1 b^2_1 \cdots b^1_m b^2_m 0 b^2_{m+1} 1 b^2_{m+2} 1 b^2_{m+3} \cdots\\
\end{aligned}
$$

This shows the injectivity is broken, the same point $(v_1, v_2)$ can be mapped to by (at least) two different $\alpha \neq \alpha^\prime$. The Z-curve $Z^{-1}$ therefore no longer defines a one-to-one mapping, and is not injective.

For a point $(v_1, v_2) \in [0,1]^2$ which can be written as a finite binary expansion (that is, a $1$ followed by an infinite string of $0$'s), there might as much as 4 pre-images in $[0,1]$:

$$
\begin{aligned}
(v_1, v_2) &= (0.b^1_1 \cdots b^1_{m_1} 1000 \cdots, 0.b^2_1 \cdots b^2_{m_2} 1000 \cdots) \\
&= (0.b^1_1 \cdots b^1_{m_1} 1000 \cdots, 0.b^2_1 \cdots b^2_{m_2} 0111 \cdots) \\
&= (0.b^1_1 \cdots b^1_{m_1} 0111 \cdots, 0.b^2_1 \cdots b^2_{m_2} 1000 \cdots) \\
&= (0.b^1_1 \cdots b^1_{m_1} 0111 \cdots, 0.b^2_1 \cdots b^2_{m_2} 0111 \cdots) \\
\end{aligned}
$$

has the following pre-images (supposing without loss of generality $m_1 \leq m_2$):

$$
\begin{aligned}
\alpha_1 &= 0.b^1_1 b^2_1 \cdots b^1_{m_1} b^2_{m_1} 1 b^2_{m_1+1} 0 b^2_{m_1+2} 0 \cdots b^2_{m_2} 010000 \cdots \\
\alpha_2 &= 0.b^1_1 b^2_1 \cdots b^1_{m_1} b^2_{m_1} 1 b^2_{m_1+1} 0 b^2_{m_1+2} 0 \cdots b^2_{m_2} 010101 \cdots \\
\alpha_3 &= 0.b^1_1 b^2_1 \cdots b^1_{m_1} b^2_{m_1} 0 b^2_{m_1+1} 1 b^2_{m_1+2} 1 \cdots b^2_{m_2} 101010 \cdots \\
\alpha_4 &= 0.b^1_1 b^2_1 \cdots b^1_{m_1} b^2_{m_1} 0 b^2_{m_1+1} 1 b^2_{m_1+2} 1 \cdots b^2_{m_2} 101111 \cdots
\end{aligned}
$$

## The set of points with multiples preimages in $[0,1]$ is dense in $[0,1]^2$

The Z-curve is not injective from $[0,1]$ into $[0,1]^2$, and furthermore, we cannot ignore the points for which it is not injective, because the set of points with multiple preimages in $[0,1]$ is dense in $[0,1]^2$. We begin the study by considering only the first coordinate $v_1$, since the same arguments apply to $v_2$.

The set $\P$ is defined as the set of $v_1$ which have multiple pre-images in $[0,1]$:

$$
\P = \{ 0.b_1 b_2 \cdots b_m 1000 \cdots \ |\ m \in \mathbb{N} \}.
$$ {#eq-P}

We wish to prove that $\P$ is dense in $[0,1]$. For this, we will prove that any point $v_1 \in [0,1]$ is either in $\P$, or is the limit point of $\P$. If $v_1 \in \P$, we are done. Else, we consider $v_1 \notin \P$.

We define an expansion of $v_1 = \sum_{l=1}^\infty b_l 2^{-l}$ (we note $b_l$ are not unique, but this does not matter for the rest of this proof). We also define the truncation $v_1^{(m)} = \sum_{l=1}^m b_l 2^{-l}$, such that $v_1^{(m)} \xrightarrow{m\to\infty} v_1$. The trick is now to define the sequence of $v_1^\prime{^{(m)}}$ as:

$$
v_1^\prime{^{(m)}} = v_1^{(m)} + \frac{1}{2^{m+1}} \in \P.
$$

Therefore, the sequence $\{v_1^\prime{^{(m)}}\}$ is a subset of $\P$, and its limit is:

$$
\lim_{m\to\infty} v_1^\prime{^{(m)}} = \lim_{m\to\infty} v_1^{(m)} + \lim_{m\to\infty} \frac{1}{2^{m+1}} = v_1.
$$

Therefore, $v_1$ is a limit point of $\P$. This concludes the proof that $\P$ is dense in $[0,1]$.

By extension, the same arguments apply to $v_2$, and the set $\P \times \P$ is dense in $[0,1]\times [0,1]$. Therefore, the set of points with multiple preimages in $[0,1]$ is dense in $[0,1]^2$.

<!-- 
For this, we adopt a counting argument and look at the scaling behavior of $\P_n$, the finite-$n$ truncations of $\P$.

Let us take $n > m$, and the finite-$n$ truncations $v_1^{(n)}$ of $v_1 \in \P$. Because of the ambiguity of representation, we need to consider two different finite-$n$ truncations, that both converge to $v_1$ as $n\to\infty$. This defines the sequence of sets $\P_n$ as:

$$
\P_n = \{ 0.\underbrace{b_1 b_2 \cdots b_m}_{m\ \text{bits}} \underbrace{1000 \cdots 0}_{n-m\ \text{bits}} \ |\ m < n \} \cup \{b_1 b_2 \cdots b_m 0111\cdots 1\ |\ m < n\}.
$$ {#eq-Pn}

At finite $n$, we can define the complement, that is, the set of non-problematic points as:

$$
\begin{aligned}
\P_n^c &= \B_n \backslash \P_n \\
\text{where}\ \B_n &= \left\{ \frac{k}{2^n} \ |\ k \in \{0, \cdots, 2^{n}-1\}\right\}
\end{aligned}
$$ {#eq-Pnc}

We now argue based on the size of $\P$ and $\P^c$. From its definition, we directly see that $|\B_n| = 2^n$. To compute $|\P_n|$, it is easier to rewrite $\P_n$ as the union of $m$ *disjoint* sets $\P_{n,m}$:

$$
\begin{aligned}
\P_n &= \bigcup_{m=1}^n \P_{n,m} \\
&= \bigcup_{m=1}^n \left\{ \frac{k}{2^m} + \frac{1}{2^{m+1}}, \frac{k}{2^m} + \frac{2^{m+1}-1}{2^{m+1}}\ |\ k \in \{0, \cdots, 2^{m} - 1\},  \right\}.
\end{aligned}
$$

Since each set $\P_{n,m}$ is of size $|\P_{n,m}| = 2 \times 2^{m}$, the size of $|\P_n|$ can be computed as:

$$
|\P_n| = \sum_{m=1}^n |\P_{n,m}| = 2 \sum_{m=1}^n 2^m = 2 (2^n - 1)
$$

uh oh... -->
<!-- 
## $Z^{-1}$ defines a space-filling curve on a subset of $[0,1]$ {#sec-zcurve-lebesgue}

### The Cantor set $\C$

the cantor set has measure 0, but has the same cardinality of the continuum

### $Z^{-1} | \C$ is a surjection

### $Z^{-1} | \C$ is continuous

## Relation with the Lebesgue space-filling curve

we would also relax and say the Z-curve is discontinuous, because lim alpha ->0.5+ Z(alpha) != lim alpha -> 0.5- Z(alpha)
-> NO ! because we linearly interpolate between the quadrants. we say the curve is continuous, in the sense that it is connected. but it is allowed to jump around !

talk about the fact we have other space filling curves, such as hilbert curve or peano curve, that are continuous and surjective [0,1] -> [0,1]^2. we use the Z-curve because it's easy to implement numerically -->