### 2023-06-21 manuscript/chapters/3_mappings.qmd

# Mapping neural fields to lower-dimensional embeddings {#sec-mappings}

::: {.content-hidden}
{{< include ../macros.tex >}}

```{python}
import matplotlib as mpl, matplotlib.pyplot as plt, matplotlib.colors as mplcolors, matplotlib.patheffects as pe, matplotlib.ticker as ticker
import numpy as np
import sys
sys.path.append('../../notebooks')  # import neurodyn
from neurodyn import *
from matplotlib_tufte import *
setup()

mpl.use('svg')

def mktrans(fig, ax):
	"""Makes figure and axes transparent"""
	fig.patch.set_facecolor('#FFFFFF00')
	ax.patch.set_facecolor('#FFFFFF00')
```
:::

In the previous chapter, we have established a simple toy model for a neural field in $p$ dimensions, and have shown how it can be embedded in $[0,1]^p$. In this chapter, we discuss in-depth what types of mappings can be used to map the $p$-dimensional space to $[0,1]$, and in particular we study the case $p=2$. Before getting into too much detail, let us give some insight as to why this problem is not trivial.

## Population dynamics: an introductory example of a 1D embedding

### Embedding a finite number of populations in $[0,1]$

Let us for a moment consider another type of model for networks of neurons, that of multiple homogeneous population dynamics, describing how $N_\text{pop}$ different populations interact with each other. These types of models emerge, for example, when networks of neurons exhibit a finite number of "neuron types". We then group neurons of the same type, and make the approximation that they have identical potentials and the same interaction weights with other neurons. As a consequence, every population represents a homogeneous population of identical neurons.

This results in writing population dynamics with the same equation as a finite network of neurons. However, the equation *already represents* the $N \to \infty$ limit. We write the potential of the populations $H_a(t)$ and the population connectivity $\tilde J_{ab}$, where $a,b \in \{1, \cdots, N_\text{pop}\}^2$ in @eq-pop-dynamics.

$$
\dot H_a(t) = -H_a(t) + \sum_{b=1}^{N_\text{pop}} \tilde J_{ab} \phi(H_b(t))
$$ {#eq-pop-dynamics}

We don't show the derivation of population dynamics here (see @GerKis14, chapter 12.3 for a detailed analysis), but we point out the implicit scaling behavior of $\tilde J_{ab}$. Letting $i$ be a neuron of population $a$ and $j$ be a neuron of population $b$, the connectivity between the two neurons is $J_{ij}$, and the population connectivity scales with the number of neurons in the presynaptic population $|b|$. Furthermore, we make the reasonable assumption that $J_{ij}$ scales as $J^0_{ij} / N$, where $J^0_{ij} \sim \bO(1)$. We then have

$$
\tilde J_{ab} = \frac{|b|}{N} J^0_{ij}.
$$

As $N\to\infty$, the ratio $\tfrac{|b|}{N}$ converges to a constant $p_b$, the probability of the population, normalized such that $\sum_{a=1}^{N_\text{pop}} p_a = 1$. It expresses the probability of uniformly sampling a neuron belonging to population $b$. In effect, $p_b$ can be interpreted as the "weight" of the population.

So far, we have not considered that the populations are embedded in some abstract space; but the notion of population weights gives rise to the obvious embedding in $[0,1]$ given by the cumulative sums of the population weights, illustrated in @fig-popdynamics-01. We define (somewhat informally) the connectivity kernel $w(\alpha, \beta) = J^0_{ab}$ for $\alpha, \beta$ in the regions corresponding to populations $a$ and $b$ respectively.[^poplocation]

[^poplocation]: Formally, we would write the cumulative sum $K_a = \sum_{b=1}^a p_b$. Then, we have $\alpha \in [K_{a-1}, K_{a})$ and $\beta \in [K_{b-1}, K_{b})$. We abusively wrote $J^0_{ab}$ since all neurons in each population are identical, therefore we can index into the connectivity $J^0_{ij}$ with the population indices $a,b$ instead of neuron indices $i,j$.

![Embedding of the population dynamics in $[0,1]$.](figures/fig-multipop.svg){width=500px #fig-popdynamics-01}

Defining the neural field $H(\alpha, t) = H_a(t)$, we can write a "neural field equation" on the $[0,1]$ space.

$$
\begin{aligned}
\partial_t H(\alpha, t) &= -H(\alpha, t) + \int_0^1 w(\alpha, \beta) \phi(H(\beta, t)) \d \beta \\
&= -H(\alpha, t) + \sum_{b=1}^{N_\text{pop}} p_b w(\alpha, b) \phi(H_b(t)) \\
\iff \partial_t H_a(t) &= -H_a(t) + \sum_{b=1}^{N_\text{pop}} p_b w(a, b) \phi(H_b(t)) \\
&= -H_a(t) + \sum_{b=1}^{N_\text{pop}} p_b J^0_{ab} \phi(H_b(t)) \\
&= -H_a(t) + \sum_{b=1}^{N_\text{pop}} \tilde J_{ab} \phi(H_b(t)) 
\end{aligned}
$$

We hereby showed (having taken many writing liberties, but this is just to give an idea) how the neural field reduces back to the original population dynamics, which was in itself already the $N\to\infty$ limit of the network. The case of multi-population dynamics therefore shows a trivial example of an embedding in $[0,1]$.

### Population dynamics with an infinite number of populations

When taken in isolation, the previous developments might sound a little bit silly, though they serve to give a feeling for the problem at hand, and a motivation for the rest of this work. For a finite number of populations $N_\text{pop}$, the population dynamics is well-defined; however, what happens as $N_\text{pop} \to \infty$ ? The core of the question lies in the fact that we can easily *enumerate* a finite number of populations, but the sum $\sum_{b=1}^{N_\text{pop}}$ must have a clearly defined interpretation when this number tends to infinity.

In light of the toy model presented in @sec-nf, we can reinterpret the density $\rho$ of the neural field equation. Every point $\vec z \in \R^p$ (let us consider $p=2$ in the following) describes an infinite population of homogeneous neurons, associated with a probability (density) $\rho(\vec z)$. Therefore, in order to write an equivalent neural field in $[0,1]$, we need a way to *enumerate* the populations. This is done using a mapping $S : [0,1]^2 \mapsto [0,1]$, where we cover the $[0,1]^2$ space by following the order assigned to the populations in the image $[0,1]$.

Additionally, the notion of population dynamics helps give another interpretation of the grid integration method given in @sec-gridmethod. When we discretize the $[0,1]^2$ space, we are effectively performing a coarse-graining approximation and simulating a finite number of populations. For every square bin localized by $(v_1, v_2)$, we are sampling the connectivity kernel $w_U((v_1, v_2), \cdot)$ and the 2D neural field $h_U((v_1, v_2), t)$ (see @fig-grid-sampling). In the rest of the chapter, we refer to these 2D bins as "square populations", and to the corresponding 1D bins obtained by the mapping $S$ as "segment populations".

![Every point of the grid effectively performs coarse-graining, and defines a point where we sample the connectivity kernel $w_U$ and the 2D neural field $h_U$.](figures/fig-smallsquare.svg){width=500px #fig-grid-sampling}

For a finite number of populations $N_\text{pop}$, applying a bijective mapping $S$ will evidently give an equivalent neural field, because the connectivity matrix is permutation-invariant, and the resulting connectivity matrix $J_{S(a),S(b)}$ will therefore give equivalent dynamics. As we increase $N_\text{pop}$ by taking finer and finer grids, the image of the grids defines a sequence of mappings $S^n: \text{grid in}\ [0,1]^2 \mapsto \text{grid in}\ [0,1]$, which we aim to study in the rest of this chapter. The main question is:

> How do we define a sequence of mappings $S^n$ such that the limiting $S : [0,1]^2 \mapsto [0,1]$ can express a 1D neural field with identical dynamics ?

## Defining the neural field in $[0,1]$

We start by writing the neural field equation in the $[0,1]$ space. 
For $\lambda: [0,1]^2 \to [0, \infty]$ a measurable function (e.g. the Lebesgue measure, which is nothing but the uniform probability distribution), we let the composition $\lambda \circ S^{-1}$ define a measure on $[0,1]$, which keeps track of how 2D surfaces are transformed to (possibly disjoint) 1D segments through the mapping $S$.

The connectivity kernel $\tilde w$ defines the interactions between the populations in the $[0,1]$ embedding, which we define relating to the kernel $w_U$, defined in @sec-mapping-cdf as:

$$
\tilde w(\alpha, \beta) = w_U(S^{-1}(\alpha), S^{-1}(\beta))
$$ {#eq-kernel-1d}

Defining the initial condition for the 1D neural field as $\tilde h(\alpha, t=0) = h_U(S^{-1}(\alpha), t=0)$, we have fully defined the equation of evolution in the one-dimensional embedding in @eq-nf-1d.

$$
\partial_t \tilde h(\alpha, t) = -\tilde h(\alpha, t) + \int_{[0,1]} \tilde w(\alpha, \beta) \phi(\tilde h(\beta, t)) \left[\lambda \circ S^{-1}\right](\d \beta)
$$ {#eq-nf-1d}

The definition of the connectivity kernel in @eq-kernel-1d might seem innocent, but the key element here is the inverse $S^{-1}$. We recall that in the toy model, the original connectivity kernel is written as $w(\vec z, \vec y) = \sum_{\mu=1}^p z_\mu \tilde \phi(y_\mu)$. This kernel is highly regular, by which we mean that, by its continuity in each of its arguments, populations nearby in the $\Rp$ will have similar input weights. Since it is also differentiable, we can write this as a Taylor expansion in the first argument:

$$
\begin{aligned}
w(\vec z + \vec \epsilon, \vec y) &= w(\vec z, \vec y) + \sum_{\mu=1}^p \epsilon_\mu \frac{\partial w}{\partial z_\mu} + \bO(\norm{\vec \epsilon}^2) \\
&= w(\vec z, \vec y) + \sum_{\mu=1}^p \epsilon_\mu \phi(y_\mu) + \bO(\norm{\vec \epsilon}^2) \\
&\approx w(\vec z, \vec y).
\end{aligned}
$$

When we do the change of variables with the CDF, we don't encounter problems because the Gaussian (inverse) CDF is differentiable and its derivative is continuous;[^inversecdf] therefore $w_U(\vec v, \vec u) = w(\textrm{CDF}^{-1}(\vec v), \textrm{CDF}^{-1}(\vec u))$ is still highly regular.

[^inversecdf]: Strictly speaking, we have the problem that $\textrm{CDF}^{-1}(v) \xrightarrow{v \to 1} \infty$, and so the derivative diverges when $v \to 1$ (and similarly, when $v \to 0$). However, this is not a problem, since these limits correspond to $z \to \pm \infty$, where the density vanishes.

For these reasons, we can see why the composition of $w_U$ with a possibly non-differentiable mapping $S$ might be problematic. Since the only thing we know about $S$ is that it is measurable and bijective, *there is no guarantee that the regularity (continuity and the differentiability) of the kernel is conserved through composition with the mapping $S$.*

## Mappings of $[0,1]^2$ to $[0,1]$

### The search for differentiable mappings

We start our search of mappings by attempting to find a mapping $S$ that is differentiable.
The easiest example of such a mapping would be a linear projection, for instance, $P(v_1, v_2) = v_1$. This mapping is obviously continuous and differentiable in both components, and surjectively maps the unit square to the unit segment.
However, the obvious problem with this approach is that linear projections are not invertible. Hence, should we nevertheless define an inverse $P^{-1}$ that computes the preimage of $\alpha \in [0,1]$, we have that:

$$
P^{-1}(\alpha) = \{(\alpha, v_2)\ |\ v_2 \in [0,1] \}.
$$

In other words, $P^{-1}(\alpha)$ defines a vertical line in the $[0,1]^2$ embedding. When we take the composition $w_U \circ P^{-1}$, the "locality" is destroyed, in the sense that the same population $\alpha$ in the 1D embedding will "sense" all the connectivity kernels on the vertical line.

A more sophisticated attempt, in the spirit of @sec-mapping-cdf, might be to consider the joint cumulative function

$$
\alpha = \mathrm{CDF}(z_1, z_2) = \int_{-\infty}^{z_1} \int_{-\infty}^{z_2} \rho(y_1, y_2) \d y_1 \d y_2.
$$

However, this again fails because the joint cumulative function is not bijective, and isolines of the CDF define continuous curves in the uniform embedding space $[0,1]^2$.[^jointcdf]

[^jointcdf]: We note to avoid any confusion, that in @sec-mapping-cdf the CDF is taken component-wise, instead of jointly. The univariate Gaussian CDF defines a bijection between $\R$ and $(0,1)$, but the joint CDF does not.

Perhaps these examples are a bit too naive, and we need to look for more complex mappings. We might ask the question of whether a diffeomorphism (a differentiable bijection with a differentiable inverse) even exists between the unit square and the unit segment. Unfortunately, results from topology give a negative answer, and even worse, it can be shown that there is no homeomorphism (a continuous bijection with continuous inverse) between the unit square and the unit segment. This is stated by Netto's theorem, which imposes continuous bijections to conserve the dimensionality between the domain and the image.

@fig-venn shows a visual representation of Netto's theorem. The 3-way intersection of the Venn diagram represents that functions which are both bijective (surjective and injective) and continuous do not exist. For the three remaining 2-way intersections, we annotate, without going into detail, the types of functions which they represent (see [@kharazishvili_strange_2017] for Cantor and Peano functions, [@sagan_space-filling_1994] for Peano and Jordan functions).

<!-- ### Mappings of $[0,1]$ into $[0,1]^2$

Since in the composition $w_U \circ S^{-1}$ we work with the inverse mapping, we contextualize our study by first considering the types of functions from $[0,1]$ into $[0,1]^2$. We have already established that there is no homeomorphism between those spaces, and need to relax one of the conditions. Since bijections are functions that are both injective and surjective, we can represent Netto's theorem as a 3-way Venn diagram in @fig-venn. -->

![Venn diagram reprensenting the properties of functions $[0,1] \mapsto [0,1]^2$ which are allowed by Netto's theorem.](figures/fig-venn.svg){width=400px #fig-venn}

<!--
* Cantor functions are bijections $[0,1] \mapsto [0,1]^2$. However, they are not continuous, and thereby break the continuity of the kernel.
<!-- Nicole: ACTUALLY NO THIS IS NOT A PROBLEM !! OUR Z-MAPPING IS NOT CONTINUOUS, BUT LOCALITY STILL SAVES US, SO THAT IT IS CONTINUOUS "ON AVERAGE"
* Jordan curves are continuous injections $[0,1] \mapsto [0,1]^2$. Since they are not surjective, these curves cannot entirely "fill" the $[0,1]^2$ space, and we will not use them.
<!-- Nicole : see my notebook about Jordan functions and their nonzero Jordan measure
* Peano curves are continuous surjections $[0,1] \mapsto [0,1]^2$. Their non-injectivity however is not a problem, since it just means that two segment populations can "refer" to the same square population (see @fig-non-injection). Therefore, Peano functions seem like good candidates for the mapping $S^{-1}$.
-->

<!-- 
Functions of $[0,1]$ into $[0,1]^2$ are extensively discussed in chapter 1 of @kharazishvili_strange_2017, and we cite two classes of such functions :

1. Cantor functions are bijections between $[0,1]$ and $[0,1]^2$. By Netto's theorem, these functions cannot be continuous.
2. Peano functions are continuous surjections from $[0,1]$ to $[0,1]^2$. They cannot be injections. For reasons that will become apparent in the following sections, Peano functions are often called space-filling curves.

In light of these results, we might be tempted to search for expressions of Cantor-type functions, since having a bijection would allow us to have a direct mapping between the square and the segment populations. Problematically, these functions are discontinuous and thereby break the continuity of the kernel.

Peano functions seem like good candidates for the mapping $S^{-1}$, and are extensively studied in the literature around space-filling curves (we refer for instance to [@sagan_space-filling_1994]). They have the nice property that they are continuous and surjective, so the segment populations can "reach" the entirety of the square populations. All things considered, non-injectivity is not a problem, since it just means that two segment populations can "refer" to the same square population (see @fig-non-injection), which does not break the regularity of the kernel,  since the two segment populations can just have their own neighbourhoods.
-->

<!-- ![Non-injectivity of $S^{-1}$ means that two segment populations can refer to the same square population.](figures/photo_2023-06-12_00-46-26.jpg){#fig-non-injection}

Nicole: removed. too confusing
-->

### Sequences of bijective mappings

#### Mappings as binary expansions {#sec-mappings-binary}

Previously, we motivated $S$ as the limit of functions $S^n$ acting on finer and finer grids. Let us consider $(v_1, v_2) \in [0,1]^2$. Given $n \in \mathbb N$, we write a (truncated) binary expansion of $v_1$ and $v_2$.

$$
\begin{aligned}
v^{(n)}_1 &= \sum_{l=1}^{n} b^1_l 2^{-l} = 0.b^1_1b^1_2\cdots b^1_{n} \iff b^1_l = \mathrm{Ind}\left\{2^{l-1}v_1 - \lfloor2^{l-1}v_1\rfloor \geq \frac{1}{2}\right\} \\
v^{(n)}_2 &= \sum_{l=1}^{n} b^2_l 2^{-l} = 0.b^2_1b^2_2\cdots b^2_{n} \iff b^2_l = \mathrm{Ind}\left\{2^{l-1}v_2 - \lfloor2^{l-1}v_2\rfloor \geq \frac{1}{2}\right\}
\end{aligned}
$$ {#eq-trunc}

@fig-recurse-binary illustrates this expansion: we locate each component bit by recursively splitting the segment in two sub-segments of equal length. At every step, taking the left or right segments defines the sequence of binary digits.
<!-- Then combine the bits of $v^{(n)}_1$ and $v^{(n)}_2$ to get the position in the $[0,1]^2$ embedding. -->

![Splitting $[0,1]$ recursively with the indicator function gives the binary bit expansion of $v_1$](figures/fig-binary.svg){width=400px #fig-recurse-binary}

We define the mapping $S^n$ that takes the $n$-bit truncations $v^{(n)}_1$ and $v^{(n)}_2$ as input, and outputs a $2n$ bit truncation $\alpha^{(n)}$ of $\alpha=S(v_1, v_2)$, where $S$ is the $n\to\infty$ pointwise limit of the sequence $S^n$ (when the sequence $S^n$ converges).

$$
\begin{aligned}
S^n : &\{0, \tfrac{1}{2^n}, \cdots, 1-\tfrac{1}{2^n}\}^2 \mapsto \{0, \tfrac{1}{4^n}, \cdots, 1-\tfrac{1}{4^n}\} \\
&(v^{(n)}_1,v^{(n)}_2)=(0.b^1_1 b^1_2 \cdots b^1_n, 0.b^2_1 b^2_2 \cdots b^2_n) \mapsto \alpha^{(n)}=0.b_1 b_2 \cdots b_{2n}
\end{aligned}
$$ {#eq-Sn}

In general, each bit $b_k$ of the image would be a binary function of the $n$ bits of $v_1$ and the $n$ bits of $v_2$, i.e. $b_k = b_k(b^1_1, b^1_2, \cdots, b^1_n, b^2_1, b^2_2, \cdots, b^2_n)$.
With this, we can formulate any mapping between sets of $4^n$ distinct elements.

The intuition behind the use of $2n$ bits for the image is that we take $2^n \times 2^n = 4^n$ squares of size $2^{-n} \times 2^{-n}$ in $[0,1]^2$, and map those to $4^n$ segments of size $4^{-n}$ in $[0,1]$. Informally, we map $n+n$ bits onto $2n$ bits.

In the following, we will restrict ourselves to mappings where the functions $b_k$ are defined in such a way that each bit of the inputs $v_1$ and $v_2$ is used only once. Therefore, for each bit $k$ of the output, there is a function $b_k(b^{\mu_k}_{l_k})$ acting on only one bit of the input, where $\mu_k \in \{1,2\}$ is the input component and $l_k \in \{1,\cdots,n\}$ the bit number corresponding to $k$. Since this function is binary, the only allowed definitions are

* $b_k(b^{\mu_k}_{l_k}) = b^{\mu_k}_{l_k}$ (identity),
* $b_k(b^{\mu_k}_{l_k}) = 1-b^{\mu_k}_{l_k}$ (inversion),
* $b_k(b^{\mu_k}_{l_k}) = 0$ (constant zero),
* $b_k(b^{\mu_k}_{l_k}) = 1$ (constant one).

The constant functions are ignored, since they effectively just restrict the output domain to a subset of $[0,1]$; and since only $2n$ bits are allowed for the output, they are a "waste" by ignoring information of the input.
Because they don't add information to the output (they just "swap" the values of the output bits at known locations), inversion functions are also not considered.

All this mathematical formalism expresses, in essence, that the mapping $S^n$ is just a *reordering* of the input bits. $S^n$ takes $4^n$ points embedded in $[0,1]^2$, and reorders them onto $4^n$ points in $[0,1]$.[^edges]

[^edges]:
  We note that in our definition, $S^n$ maps $(0., 0.)$ onto $0$. Numerically, we encounter the problem that the inverse of the Gaussian CDF, used to compute the connectivity, goes to infinity when $v_\mu$ (one of the components of $(v_1, v_2) = S^{-1}(\alpha)$) goes to $0$ (or $1$). To avoid this, we offset the squares (by $2^{-n-1}$ along each dimension), so that geometrically their position is not indexed by their lower-left corner, but rather their center (this is represented in @fig-grid-sampling). Analytically, this is not a problem, because the kernel defined in @eq-hu-wu diverges only linearly, compared to the exponential cutoff of the Gaussian distribution, so that the integrand still vanishes at infinity.

Therefore, there is a one-to-one mapping between the points on the square and the points on the segment, and the mapping $S^n$ is bijective. In the following, we will, for ease of notation and when it is clear in the context, drop the index $n$ when writing $S^n$. This is often justified (particularly with the Z-mapping and Column mapping introduced later), because the finite-$n$ truncations $(v_1^{(n)}, v_2^{(n)})$ result in finite truncations of $S$, therefore $S(v_1^{(n)}, v_2^{(n)}) = S^n(v_1^{(n)}, v_2^{(n)})$.

<!-- These simplifications might seem too restrictive, however, -->
We will show in the rest of this thesis that this simple formulation is flexible enough to express mappings that have good properties such as "locality", while being simple enough to help build intuitions.

<!-- #### A geometric view relating to space-filling curves
Let's not mention space-filling curves, because the Column mapping is not space-filling (no continuity 1D -> 2D), and because we would need a whole lot more formalism, see @Sagan1994
 -->
#### A geometric view relating to curves in $[0,1]^2$

With this formalism in mind, we give a geometric interpretation to the mappings $S^n$. The core idea is that the populations in $[0,1]$ have a clear ordering to them, and we can enumerate them simply by their position in the embedding. Since we constructed $S^n$ to be bijective, this enumeration traces a path in the corresponding populations of the $[0,1]^2$ embedding.

We illustrate this by introducing the "column" mapping.

$$
\alpha^{(n)} = C(v^{(n)}_1,v^{(n)}_2) = 0.b^1_1 b^1_2 \cdots b^1_n b^2_1 b^2_2 \cdots b^2_n
$$ {#eq-column}

The intuition behind this mapping is that we enumerate the populations on a grid column by column, letting the index of the rows vary the fastest. We visualize this mapping in @fig-col, by coloring the square populations by the position of the corresponding $[0,1]$ population, and additionally drawing lines between the points to help guide the eye.

```{python}
#| label: fig-col
#| fig-cap: Column mapping for $n=2$. Numbers represent the ordering of the populations in the $[0,1]$ embedding, neighbouring populations are connected by lines.
#| echo: false
def plot_trace(mapping_cls, nrec, do_trace: bool = True, do_numbers: bool = True):
	F = ColumnMapping.new_nrec(nrec).inverse_samples(centered=True)
	mapping = mapping_cls.new_nrec(nrec)
	alpha = mapping(F)

	fig, ax = plt.subplots(facecolor='#FFFFFF00')
	im = ax.imshow(alpha.reshape((2**nrec, 2**nrec)).T, origin='lower', vmin=0, vmax=1, extent=(0,1,0,1), interpolation='none')
	mapping_letter = { ZMapping: 'Z', ColumnMapping: 'C', AntiZMapping: 'A' }.get(mapping_cls, 'S')
	plt.colorbar(im, ax=ax, label=f'$\\alpha={mapping_letter}^{{n}}(v_1, v_2)$')
	if do_trace:
		# plot the trace
		ax.plot(*F[np.argsort(alpha)].T)
		# hide trace at numbering location
		ax.scatter(*F.T, s=250, c=alpha, cmap='viridis', vmin=0, vmax=1, zorder=10)  
	# add the bin numbers
	if do_numbers:
		for xy, i in zip(F[np.argsort(alpha)], range(1, 4**nrec+1)):
			ax.annotate(i, xy=xy, va='center_baseline', ha='center', zorder=20, color='white', fontweight='bold',
				path_effects=[pe.withStroke(linewidth=1, foreground='black')])
	ax.set_aspect('equal')
	ax.set_xlabel('$v_1$'); ax.set_ylabel('$v_2$')
	ax.set_xlim((0,1)); ax.set_ylim((0,1))
	ax.xaxis.set_major_locator(ticker.MultipleLocator(1/2**2))
	ax.yaxis.set_major_locator(ticker.MultipleLocator(1/2**2))
	# for x in np.linspace(0, 1, 2**nrec+1): ax.axvline(x, ymin=0, ymax=1, color='tab:gray', linewidth=0.5, zorder=0, clip_on=False)
	# for y in np.linspace(0, 1, 2**nrec+1): ax.axhline(y, xmin=0, xmax=1, color='tab:gray', linewidth=0.5, zorder=0, clip_on=False)
	breathe()
	return fig, ax

fig, ax = plot_trace(ColumnMapping, nrec=2)
display(fig)
```

For finite $n$, we write the inverse of the column mapping:

$$
\begin{aligned}
v^{(n)}_1 = C^{-1}_1(\alpha^{(n)}) &= 0.b_1 b_2 \cdots b_n = \sum_{k=1}^n b_{k} 2^{-k} \\
v^{(n)}_2 = C^{-1}_2(\alpha^{(n)}) &= 0.b_{n+1} b_{n+2} \cdots b_{2n} = \sum_{k=1}^n b_{n+k} 2^{-k}.
\end{aligned}
$$ {#eq-column-inverse}

This way of enumerating the square populations might be familiar to people who have worked with numerical arrays. In such (contiguous) arrays, the values are laid out in memory in "row-major" order (C-style) or "column-major" order (Fortran-style); and we can go from the 1D memory representation to a 2D matrix representation by using "reshape" operations. Effectively, this is how this mapping is implemented in code, using [`numpy.ravel_multi_index`](https://numpy.org/doc/stable/reference/generated/numpy.ravel_multi_index.html) (2D to 1D) and [`numpy.unravel_index`](https://numpy.org/doc/stable/reference/generated/numpy.unravel_index.html) (1D to 2D).

#### "Local" curves {#sec-local}

Research in computer science and efficient data structures has given rise to other ways of storing 2-dimensional information in a 1-dimensional memory.
A problem in GPU computing is that of texture locality: sampling textures often involves reading data corresponding to a small 2D region, and therefore to improve speed (minimize cache misses), the 2D data should be packed close together in the 1D memory.
One way of ordering the 2D texture data in the 1D memory is by using the Z-order curve (also known as "Morton mapping", and similar to the "Lebesgue curve", see @sagan_space-filling_1994, Chapter 6). In this way, points close in 1D *tend* to be tightly packed in 2D. We define the Z-mapping in the following way:[^zcurve]

[^zcurve]: The Z-order curve is often defined in a transpose manner, putting first the bits of $v_2$ (vertical component) then the bits of $v_1$ (horizontal component), instead of $v_1$ then $v_2$. By doing this, we get the "upright" Z shapes, as [shown here](https://en.wikipedia.org/wiki/Z-order_curve#/media/File:Four-level_Z.svg).

$$
\alpha^{(n)} = Z(v^{(n)}_1,v^{(n)}_2) = 0.b^1_1 b^2_1 b^1_2 b^2_2 \cdots b^1_n b^2_n = \sum_{k=1}^{n} b^1_k 2^{1-2k} + b^2_k 2^{-2k}
$$ {#eq-z}

And its inverse is:

$$
\begin{aligned}
v^{(n)}_1 = Z^{-1}_1(\alpha^{(n)}) &= 0.b_1 b_3 \cdots b_{2n-1} = \sum_{k=1}^n b_{2k-1} 2^{-k} \\
v^{(n)}_2 = Z^{-1}_2(\alpha^{(n)}) &= 0.b_2 b_4 \cdots b_{2n} = \sum_{k=1}^n b_{2k} 2^{-k}.
\end{aligned}
$$ {#eq-z-inverse}

@fig-z is a visualization of the Z-mapping when $n=2$, and shows how it is composed of the "Z" shapes giving it its name.

```{python}
#| label: fig-z
#| fig-cap: Z-mapping for $n=2$. Numbers represent the ordering of the populations in the $[0,1]$ embedding, neighbouring populations are connected by lines.
#| echo: false
fig, ax = plot_trace(ZMapping, nrec=2)
display(fig)
```

To illustrate this notion of locality, we take a few consecutive populations in 1D and show where the corresponding populations fall in the 2D embedding. This is done in @fig-locality, in which we see that the Z-mapping seems to conserve locality from 1D to 2D: populations close in 1D seem to be close in 2D.

We also see that the Column mapping is not local in this sense, since there is a large variation along the vertical direction, which seems to always be "of order one". This can be explained by looking at the expression for $C^{-1}_2(\alpha^{(n)}) = 0.b_{n+1} b_{n+2} \cdots b_{2n}$: small variations of the order of $2^{-n}$ in the value of $\alpha$ always result in variations of order one in $C^{-1}_2$.

```{python}
#| label: fig-locality
#| fig-cap: Locality of the mappings. The 2D populations corresponding to three small segments (each of length $\tfrac{1}{16}$) of 1D populations are shown. The figures are generated with $n=6$.
#| echo: false
##| fig-subcap:
##|   - Z mapping
##|   - Column mapping
##| layout-ncol: 2
fig, axes = plt.subplot_mosaic('aa\nbc', facecolor='#FFFFFF00', figsize=(6, 4), height_ratios=[1, 5], subplot_kw=dict(facecolor='#FFFFFF00'))
nrec = 6

cmap = mpl.colormaps['viridis']
cmap.set_bad(color='#FFFFFF00')

alpha = np.linspace(0, 1, 4**nrec+1)
mask1d = (0.2 <= alpha) & (alpha < 0.2+1/4**2)
mask1d |= (0.8 <= alpha) & (alpha < 0.8+1/4**2)
mask1d |= (0.5 <= alpha) & (alpha < 0.5+1/4**2)
alpha[~mask1d] = np.nan
axes['a'].imshow(alpha[None, :], vmin=0, vmax=1, extent=((0,1,0,0.2)), interpolation='none', cmap=cmap)
axes['a'].set_aspect('auto')
axes['a'].spines.left.set_visible(False)
axes['a'].yaxis.set_visible(False)
axes['a'].set_title('Populations in 1D', fontsize='medium')
axes['a'].set_xlabel('$\\alpha$')
breathe(axes['a'])

for mapping_cls, ax in zip((ZMapping, ColumnMapping), (axes['b'], axes['c'])):
	F = ColumnMapping.new_nrec(nrec).inverse_samples(centered=True)
	mapping = mapping_cls.new_nrec(nrec)
	alpha = mapping(F)
	mask1d = (0.2 <= alpha) & (alpha < 0.2+1/4**2)
	mask1d |= (0.8 <= alpha) & (alpha < 0.8+1/4**2)
	mask1d |= (0.5 <= alpha) & (alpha < 0.5+1/4**2)
	alpha[~mask1d] = np.nan
	ax.imshow(alpha.reshape((2**nrec, 2**nrec)).T, origin='lower', vmin=0, vmax=1, extent=(0,1,0,1), interpolation='none', cmap=cmap)
	ax.set_xlabel('$v_1=S_1^{{-1}}(\\alpha)$'); ax.set_ylabel('$v_2=S_2^{{-1}}(\\alpha)$')
	ax.set_xlim((0,1)); ax.set_ylim((0,1))
	ax.xaxis.set_major_locator(ticker.MultipleLocator(1/2**2))
	ax.yaxis.set_major_locator(ticker.MultipleLocator(1/2**2))
	ax.set_title(f'To 2D with {mapping_cls.__name__}', fontsize='medium')
	breathe(ax)

display(fig)
```

Note that, in some sense, the Column mapping is local, from 2D to 1D. This can be seen from the definition of $\alpha^{(n)}=C(v^{(n)}_1,v^{(n)}_2) = 0.b^1_1 b^1_2 \cdots b^1_n b^2_1 b^2_2 \cdots b^2_n$: small variations of order $2^{-n}$ in $v^{(n)}_1$ result in small variations also of order $2^{-n}$ in $\alpha^{(n)}$. Small variations of order $2^{-n}$ in $v^{(n)}_1$ result in even smaller variations also of order $2^{-2n}$ in $\alpha^{(n)}$ ! We therefore have two notions of locality: from 1D to 2D, and from 2D to 1D; and the Column mapping shows that they are not equivalent.

We argue that in our case only the notion of locality from 1D to 2D matters. In essence, the dynamics in the 2D embedding are already known, and we ask the question of whether it is possible to write an equivalent neural field in 1D.
<!-- (in the sense that it has identical latent trajectories) -->
Without repeating the intuition given in @sec-intro-simnf, if neighbouring populations in 1D have similar potentials, then this allows us to write a neural field equation in $[0,1]$. Therefore, we would like to have the property that populations close in 1D are (inversely) mapped to populations close in 2D, which already have similar potentials.

<!-- note for me : find a formulation for the recursive local mapping. it is not trivial, since it has

```
n=1
0 0 -> 00
1 0 -> 01
1 1 -> 10
0 1 -> 11

n=2
00 00 -> 0000
01 00 -> 0001
01 01 -> 0010
00 01 -> 0011
10 00 -> 0100
11 00 -> 0101
11 01 -> 0110
10 01 -> 0111
10 10 -> 1000
11 10 -> 1001
11 11 -> 1010
10 11 -> 1011
00 10 -> 1100
01 10 -> 1101
01 11 -> 1110
00 11 -> 1111
```

this might be nice as an appendix, to show that our formulation is very simplistic, and more complex formulations can also have nice properties

this also shows our proof is "incomplete", because it relies on the fact we have a permutation of the bits, although it is not strictly necessary. we have a sufficient, but not necessary proof

also : this might be a type of peano curve or smth. see @Sagan1994

https://www.cut-the-knot.org/Curriculum/Geometry/LebesgueCurve.shtml
-->

### Sequences of mappings and their limit

#### Pointwise convergence of mappings

We now discuss what happens to the bijective mappings when $n\to\infty$. For this, let us consider again the "random mapping" considered in the introduction @sec-intro-simnf. We discussed how this hypothetical mapping maps each position in $[0,1]^2$ into a "random position" in $[0,1]$. A finite-$n$ formulation would be that the mapping corresponds to a *random permutation* of the bits, therefore, its limit is obviously not well-defined.
<!-- It is unclear how to formally write this mapping in a way that is well-defined for a given $n$, since by definition, it is random. -->
For a given position of the input $(v_1,v_2)$, we would like that the image is stable in the $n\to\infty$ limit.

$$
S^{n}(v_1, v_2) \xrightarrow{n\to\infty} \alpha =: S(v_1, v_2)
$$

This is the condition of "pointwise convergence" of the mappings.
Let us introduce another mapping, which will help to give a better intuition of what pointwise convergence means in our context. We define, for finite $n$, the "anti-Z" mapping and its inverse:
<!-- We might be tempted to write that any mapping for which we can write the binary expansion $S^n$ is pointwise convergent, however, we show by the example of the "anti-Z" mapping that this is not the case. Let us define for finite $n$ -->

$$
\begin{aligned}
\alpha^{(n)} &= A(v^{(n)}_1,v^{(n)}_2) = 0.b^1_n b^2_n b^1_{n-1} b^2_{n-1} \cdots b^1_1 b^2_1 = \sum_{k=1}^{n} b^1_{n+1-k} 2^{1-2k} + b^2_{n+1-k} 2^{-2k} \\
\iff& \\
v^{(n)}_1 &= A^{-1}_1(\alpha^{(n)}) = 0.b_{2n-1} b_{2n-3} \cdots b_{1} = \sum_{k=1}^n b_{2(n-k)-1} 2^{-k} \\
v^{(n)}_2 &= A^{-1}_2(\alpha^{(n)}) = 0.b_{2n} b_{2n-2} \cdots b_{2} = \sum_{k=1}^n b_{2(n+1-k)} 2^{-k}.
\end{aligned}
$$

```{python}
#| label: fig-anti-z
#| fig-cap: Anti-Z mapping for $n=2,3,4$. Numbers represent the ordering of the populations in the $[0,1]$ embedding. We don't plot the trace, because it has many crossovers and overlaps which makes the plot hard to read.
#| fig-subcap: 
#|   - $n=2$
#|   - $n=3$
#|   - $n=4$
#| layout-ncol: 3
#| echo: false
fig, ax = plot_trace(AntiZMapping, nrec=2, do_trace=False)
display(fig)
fig, ax = plot_trace(AntiZMapping, nrec=3, do_trace=False)
display(fig)
fig, ax = plot_trace(AntiZMapping, nrec=4, do_trace=False, do_numbers=False)
display(fig)
```

@fig-anti-z gives some insight into the problem of this mapping. For a fixed input $(v_1,v_2)$, we see that the output point position jumps around when $n$ changes, and that these jumps seem to always be of order 1. The intuition behind pointwise convergence is that as $n$ increases, the image of a point by $S^n$ is continually refined.

Consider how $\alpha^{(n)}$ fluctuates as the precision of the input increases with $n \to \infty$. As we refine the input with fluctuations of order $2^{-n}$ (we change only the bits that come after $b^1_n$ and $b^2_n$), the output $\alpha^{(n)} = b^1_n 4^{-1} + b^2_n 4^{-2} + b^1_{n-1} 4^{-3} + b^2_{n-1} 4^{-4} \cdots$ fluctuates (at least) with amplitude $\sim 4^{-2}$, independantly of $n$. Therefore the output never converges, and the anti-Z mapping is not pointwise convergent.

In our setting, how can we guarantee that the mappings are pointwise convergent? We need to show that small fluctuations in the input result in small fluctuations in the output. The binary expansion allows us to argue that for the sequence of mappings to be pointwise convergent, we need that the "least significant bits" (LSB, the bits to the right of the binary expansion) of the input are mapped to the LSB of the output. Conversely, the "most significant bits" (MSB, the bits to the left of the binary expansion) should be mapped to the MSB of the output.

We illustrate this with the Z-mapping, written in @eq-z. Corrections of order $2^{-n}$ in $v_1$ and $v_2$ induce corrections of order $2^{1-2n}$ and $2^{-2n}$ in $\alpha$ respectively. Therefore, it is easy to see that the Z-mapping is pointwise convergent as $n\to\infty$. Similarly, the Column mapping in @eq-column is also pointwise convergent, since as discussed @sec-local it has a locality of 2D to 1D.

#### Bijectivity in the $n\to\infty$ limit {#sec-ninfty-bijection}

However, the behavior of the Column mapping and that of the Z-mapping are very different in the $n\to\infty$ limit. Up until now, we have avoided this issue by considering the finite-$n$ approximation of the mappings, which by construction is numerically bijective (that is, bijective on the discretization of $[0,1]^2$ and $[0,1]$ corresponding to finite $n$). We illustrate the infinite-$n$ (numerically, large $n$) limit of these two mappings in @fig-n-to-infty.

```{python}
#| label: fig-n-to-infty
#| fig-cap: The $n\to\infty$ limit of mappings.
#| fig-subcap:
#|   - Z-mapping
#|   - Column mapping
#| layout-ncol: 2
#| echo: false
fig, ax = plot_trace(ZMapping, nrec=6, do_trace=False, do_numbers=False)
display(fig)
fig, ax = plot_trace(ColumnMapping, nrec=6, do_trace=False, do_numbers=False)
display(fig)
```

We see that the Column mapping converges to a projection on $v_1$, in other words, $\alpha=C(v_1, v_2) = v_1$, showing that despite every $C^{n}$ being numerically bijective, the limit of the sequence is not bijective. This becomes clear when looking at the binary expansion of $C$:

$$
\begin{aligned}
C(v_1, v_2) &= \lim_{n\to\infty} C^{(n)}(v_1,v_2)\\
&= \lim_{n\to\infty} 0.b^1_1 b^1_2 \cdots b^1_n b^2_1 b^2_2 \cdots b^2_n \\
&= \lim_{n\to\infty} 0.b^1_1 b^1_2 \cdots \\
&= v_1
\end{aligned}
$$

The limit is not bijective, because the bits of $v_2$ are "lost" in the $n\to\infty$ limit, and the inverse $C_2^{-1}(\alpha)$ is not well-defined, since it "takes bits between $\infty$ and $2\infty$".

The Z-mapping seems to converge to a fractal structure. Numerically, there seems to be an infinite number of discontinuities, but the sizes of these discontinuities are, for the most part, small, so that *on average* they vanish as $n\to\infty$. We won't go further into detail about this, but this gives some intuition for the arguments presented in @sec-locality.

We finish this discussion on the limit of the sequence of mappings by asking if there are other mappings similar to the Z-mapping, in the sense that they are pointwise convergent and the limit is bijective. In an informal way, we established two conditions for this in the context of the mappings described in @sec-mappings-binary:

<!--
[^zmapping-notbijective]:
  We refer to @Sagan1994 for an in-depth study, and in particular chapter 6, in which the author shows that the Z-curve converges to the Lebesgue curve.<br> 
  <!-- which is a Peano-type function (continuous and surjective from $[0,1]$ into $[0,1]^2$).<br> 
  Despite what can often be found in online resources, the Z-curve is *not* a bijection.
  <!-- because the added injectivity would violate Netto's theorem. 
  For every point in $[0,1]^2$, there are at most 4 distinct preimages in $[0,1]$. The reason for this is the ambiguity in the writing of binary expansions, and in @sec-zcurve-noninjective we give proof that the Z-curve is not injective in the $n\to\infty$ limit.<br>
  For the following, we will consider the Z-curve to be bijective by disallowing expansions of the type $0.1000\cdots=0.0111\cdots$.
  <!-- We note that in @Sagan1994 many other Peano curves (continuous surjections $[0,1]\mapsto [0,1]^2$) are presented (albeit more difficult to implement numerically). 
  <!-- Nicole: I don't say that the Lebesgue curve is continuous, because that opens up a whole other can of worms that I don't want to go into 
<!-- Nicole: OK, after discussion with valentin, our definition of the Z-mapping is not exactly the same as Lebesgue curve.
  Z-mapping might seem to approximate the Lebesgue curve, but it is not the same.
  Lebesgue is defined on the Cantor set (the cantor set avoids the discontinuities, e.g. around alpha=0.5), then extended linearly, thus continuous
  Our Z-mapping is defined through the bits OBTAINED FROM THE INDICATOR FUNCTIONS; and so it works on NON-AMBIGUOUS binary representations. Thus, our Z-mapping is a bijection.
  Our Z-mapping is also not continuous, consider alpha=0.5. So we don't have a problem with Netto's theorem.
-->

1. condition of pointwise convergence: the LSB of the input must be mapped to the LSB of the output,
2. condition of "no information loss": the MSB of the input must be mapped to the MSB of the output.

These two conditions do not leave a lot of flexibility in the design of new mappings. We can evidently define mappings where we interleave the bits in a different order, such as $0.b^2_1 b^1_1 b^2_2 b^1_2 \cdots$; or mappings where we reorder the $k$ bits of the input, as in $0.b^1_2 b^2_2 b^1_1 b^2_1 b^1_4 b^2_4 b^1_3 b^2_3 \cdots$, but the conditions restrict the reordering to be permutations of neighbouring bits. In essence, these mappings are not fundamentally different, and the Z-mapping can be seen as a "standard example" of the pointwise convergent mappings with bijective limits, which can be expressed as the permutation of binary digits.

<!-- ### Visualizing of the mappings of the dynamic 2D neural field

It is time for a little treat. Before we argue about how to simulate the dynamics of the 1D neural field, let us simulate the dynamics of the 2D neural field, then at each timestep, we map the 2D neural field to the 1D embedding. This will give us a reference of how the 1D neural field is supposed to look under the condition that its dynamics are identical.

![(TODO) Animation showing the image of the dynamics 2D neural field through different mappings $S : [0,1]^2 \mapsto [0,1]$](figures/photo_2023-06-10_17-21-19.jpg) -->

## Coarse-graining the neural field in $[0,1]$ {#sec-coarse-graining}

### Naive simulations of the neural field on $[0,1]$

In the continuation of the simulation methods established in the @sec-simnf, we adopt a grid discretization scheme to simulate the neural field on $[0,1]$, as the grid naturally yields square populations, which can directly be used with the mappings introduced previously.[^note-mc]

[^note-mc]: In order to simulate the mappings using Monte-Carlo schemes, binning methods can be applied. The intuition behind this is that sampled populations inside a small square of size $2^{-n} \times 2^{-n}$ give rise to the notion of an "average connectivity kernel", and in our case the notion of "mean patterns". More details are given in @sec-mean-patterns.

We now address the elephant in the room: how do we numerically simulate a neural field defined on -- what appears to be -- a fractal curve? Naively, we could just proceed as follows:

1. Create a grid of $2^n \times 2^n = 4^n$ squares of size $2^{-n} \times 2^{-n}$ in the $[0,1]\times[0,1]$ embedding. Each square is located by a position $\vec{v_i} = (v_{1,i},v_{2,i}),\,i\in\{1,\cdots\,2^n\}$.
2. Compute a discretized version of the 1D connectivity kernel
   $$
   \begin{aligned}
   \tilde w(\alpha_i, \beta_j) &= \tilde J_{\alpha_i,\beta_j}\\
   &= \tilde J_{S(i),S(j)}\\
   &\stackrel{\text{def}}{=} J_{ij}\\
   &= w_U(\vec{v_i}, \vec{v_j})
   \end{aligned}
   $$
   for every pair of square locations (where we denote for simplicity of notation $S(i)=S(\vec{v_i})$).
3. Simulate the neural field in $[0,1]$ on the corresponding discretized space, with a mesh size $4^{-n}$.

The problem with this approach is that since the mappings are just *permutations* of the square populations, the resulting connectivity matrix $\tilde J_{\alpha_i,\beta_j}$ is just a permutation of the rows and columns of $J_{ij}$, and will therefore trivially lead to identical neural field dynamics.

### Coarse-graining the one-dimensional neural field

In some sense, we would like to "take into account" the fact that a mapping of the square populations has been performed. Simulations of the neural field equation on $[0,1]^2$ are done by using $4^n$ square populations on a grid, but, as long as we have $4^n$ corresponding segment populations in $[0,1]$, the dynamics will trivially be the same, *independantly of the mapping* $S^n$. Therefore, we need a way to reduce the number of segment populations in $[0,1]$.

We do this using "coarse-graining", in which we average $2^n$ consecutive segment populations, each of length $4^{-n}$, in order to obtain $2^n$ coarse-grained segment populations, each of length $2^{-n}$. This operation is illustrated in @fig-averaging.

![Averaging $2^n$ of the $4^n$ segments populations of $S^n$ results in the $2^n$ segments populations](figures/fig-cg.svg){width=400px #fig-averaging}

Without coarse-graining, the mesh size along each spatial dimension is $2^{-n}$ in 2D, but directly applying the mapping results instead in a mesh size of $4^{-n}$ in 1D, which is much finer than the $2^{-n}$ per dimension. However, after coarse-graining, the effective mesh size in 1D becomes $2^{-n}$, which matches the 2D mesh size. Therefore, we are are not compensating the reduction of dimensionality by an increase in the numerical resolution.

<!-- Let us consider what happens to the mapping when $n$ is large, and in particular how $S^{n/2}$ and $S^{n}$ behave. Since the mappings are pointwise convergent and we are considering large $n$, we argue that both $S^{n/2}$ and $S^{n}$ are good approximations of the true mapping $S$. More importantly, $S^{n/2}$ can be considered as an approximation of $S^{n}$. This is interesting from a numerical standpoint since it is feasible to compute $S^{n}$, and the approximation $S^{n/2}$ can be obtained by "coarse-graining" $S^n$. Another way of thinking about this is that $S^n$ is considered to be the "mathematically exact" fractal structure and $S^{n/2}$ the numerical approximation. The way we go from $S^n$ to $S^{n/2}$ is by averaging the segment populations by batches of $2^n$, as schematized in @fig-averaging. -->

A geometrical intuition we can give for the coarse-graining is how the averaging of the segment populations leads to sampling the connectivity kernel at "average" positions in the $[0,1]^2$ embedding. In @fig-averaging-geom, we colour-code the square populations corresponding to the binned segment populations before coarse-graining.

```{python}
#| label: fig-averaging-geom
#| fig-cap: Square populations corresponding to the segment populations before coarse-graining are colored identically when they are the same bin of 1D length $2^{-n}$. Here we show $n=3$.
#| fig-subcap:
#|   - Z-mapping
#|   - Column mapping
#| layout-ncol: 2
#| echo: false

from itertools import cycle

def plot_averaging_geom(mapping_cls, nrec=3):
	mapping = mapping_cls.new_nrec(nrec)
	F = mapping.inverse_samples(centered=True)
	F_ = F.reshape((2**nrec, 2**nrec, 2))  # bins before coarse-graining
	values = mapping(F)

	fig, ax = plt.subplots()
	ax.plot(*F.T, color='gray')
	colors = cycle(['#750087','#001997','#06bd01','#ffff00','#fe941e','#ff1e26','#fcaec5','#74d4ea','#603814','#010101','#fbf7f4'])
	for i in range(2**nrec): ax.scatter(*F_[i, :, :].T, zorder=10, color=next(colors), ec='k')
	ax.set(
		xlim=(0,1), ylim=(0,1),
		xlabel='$v_1$', ylabel='$v_2$', aspect='equal',
		xticks=[0, 0.25, 0.5, 0.75, 1], yticks=[0, 0.25, 0.5, 0.75, 1])
	mktrans(fig, ax)
	breathe(ax)
	return fig, ax

fig, ax = plot_averaging_geom(ZMapping, nrec=3)
display(fig)

fig, ax = plot_averaging_geom(ColumnMapping, nrec=3)
display(fig)
```

In summary, coarse-graining allows us to compare the dynamics of a neural field in $[0,1]^2$ simulated using $2^n \times 2^n = 4^n$ square populations; with the dynamics of a neural field in $[0,1]$ simulated using $2^n$ segment populations.

<!-- One question that may arise is why we take $n/2$ instead of $n-1$. The reason for this is that we want the number of averaged segment populations to scale with $n$. If we had chosen $S^{n-1}$, then the number of segment populations is related via

$$
N_\text{pop}(n-1) = 4^{n-1} = 4^{n} / 4 = N_\text{pop}(n) / 4.
$$

In other words, we would be averaging a constant number (4 in this case) of segment populations, independently of $n$. Choosing a fraction of $n$ for the approximation then allows us to ensure that we average $2^n$ populations, since

$$
N_\text{pop}(n/2) = 4^{n/2} = 4^n / 4^{n/2} = 4^{n} / 2^{n} = N_\text{pop}(n) / 2^n.
$$

Furthermore, this number of averaged populations still vanishes, since

$$
N_\text{pop}(n/2) / N_\text{pop}(n) = 2^{-n} \xrightarrow{n\to\infty} 0.
$$

For Nicole: This is adressed by the argument of mesh size !!
-->

## Simulations of neural fields in 3D, 2D and 1D

In this section, we show numerical simulations of the neural field in $[0,1]$, obtained from applying mappings on higher-dimensional neural fields in $[0,1]^p$. We see that the coarse-graining procedure works when applied from $[0,1]^2$ to $[0,1]$. Then, we demonstrate that it can be applied iteratively, to map from $[0,1]^3$ to $[0,1]^2$, and from $[0,1]^2$ to $[0,1]$.

### Practical aspects of implementing the simulations of the one-dimensional neural field

We give a quick overview of the numerical details that go into implementing the methods presented above. Since the mappings (Z-mapping, Column mapping) can be written as binary expansions, they naturally lend themselves to numerical approximations (see [`ZMapping`](https://github.com/ninivert/lcnthesis/blob/master/notebooks/neurodyn/_mapping.py#L513), [`ColumnMapping`](https://github.com/ninivert/lcnthesis/blob/master/notebooks/neurodyn/_mapping.py#L397)). Furthermore, this is made especially easy since the mappings we defined are simply permutations of bits (in the $n < \infty$ case), and so we can write a one-to-one correspondence between unsigned integers.

We have implemented the mappings as all inheriting from the same abstract class, and each mapping simply has to define a method `indices_to_indices2d(indices: array[int, 1]) -> array[int, 2]` to go from 1D to 2D, and a method `indices2d_to_indices(indices2d: array[int, 2]) -> array[int, 1]` to go from 2D to 1D.[^indices] The abstract class simply defines additional quality-of-life functionality, which makes use of these two methods.

[^indices]: In the source code, only the method `indices(v12: array[float, 2]) -> array[int, 1]` is defined. This is because some mappings are implemented (for instance, the [`RecursiveLocalMapping`](https://github.com/ninivert/lcnthesis/blob/master/notebooks/neurodyn/_mapping.py#L244), that we don't discuss here), that don't lend themselves to the simple permutation of bits formulation.

Finally, the coarse-graining is easily implemented as the combination of a reshape operation (to bring all consecutive sequences of $2^n$ segment populations together), followed by averaging (see the [implementation here](https://github.com/ninivert/lcnthesis/blob/master/notebooks/neurodyn/_rnn.py#L186))

### Application of $S$: from $[0,1]^2$ to $[0,1]$

We start with the cycling neural field presented in @sec-cycling-nf. The kernel $\tilde w$ of the one-dimensional neural field is numerically approximated using the coarse-graining procedure of different mappings. Simulations are presented in @fig-cg-2d-1d (we again set $\delta=6$ and the initial condition $h(z_1, z_2, t < 0) = z_1$), showing the activity levels, recurrent currents, and the latent trajectory $(\kappa_1(t), \kappa_2(t))$. In order, we discuss:

* The random mapping can be numerically evaluated as a trivial case of a mapping that does not have locality. The coarse-graining averages out all the structure of the activity, and the recurrent currents are zero, because the contributions from excitatory and inhibitory populations cancel out. This results in the erasure of the cycling behavior in the latent trajectory.
* The Column mapping shows the projection of the initial condition onto the $z_1$ axis, yielding the sigmoid in the $[0,1]$ embedding. Again, the current currents cancel out and yield zero, and the latent trajectory quickly decays towards $(\kappa_1, \kappa_2) = (0,0)$.
* The Z-mapping shows that despite the fractal structure of the activity, there is still regularity in the $[0,1]$ embedding, resulting in non-zero recurrent currents. In consequence, the latent trajectory shows a clear cycling behavior, that closely matches that of the 2D neural field.

<!-- TODONE : we run simulations and compare latent trajectories. first show only column and z-mapping. add quick intro to diagonal, col/row, reclocal, mention cannot be expressed by bit permutation. diagonal, col/row do not work, despite the non-zero field and recurrent currents. reclocal does work. intuition is that the recursive approach (zmapping, reclocal) seems to give locality

Nicole: we don't show the other mappings, because they are not as nice examples, and essentially they work the same
-->

::: {.content-visible when-format="html"}

:::::: {#fig-cg-2d-1d}

![Neural field in the 2D embedding](figures/coarse_graining_p=2_original.mp4){#fig-cg-2d-1d-original}

![Random mapping followed by coarse-graining](figures/coarse_graining_p=2_random.mp4){#fig-cg-2d-1d-random}

![Column mapping followed by coarse-graining](figures/coarse_graining_p=2_column.mp4){#fig-cg-2d-1d-column}

![Z-mapping followed by coarse-graining](figures/coarse_graining_p=2_z.mp4){#fig-cg-2d-1d-z}

The cycling 2D neural field ($4^8$ square populations) is mapped to $[0,1]$ using different mappings, then coarse-grained (resulting in $2^8$ segment populations).
::::::

:::

::: {.content-visible when-format="pdf"}
```{python}
#| label: fig-cg-2d-1d
#| layout-ncol: 2
#| layout-nrow: 4
#| fig-cap: The cycling 2D neural field ($4^8$ square populations) is mapped to $[0,1]$ using different mappings, then coarse-grained (resulting in $2^8$ segment populations).
#| fig-subcap:
#|   - Neural field in the 2D embedding. First frame, animation [available here](https://ninivert.github.io/lcnthesis/chapters/figures/coarse_graining_p=2_original.mp4).
#|   - Neural field in the 2D embedding. Last frame, animation [available here](https://ninivert.github.io/lcnthesis/chapters/figures/coarse_graining_p=2_original.mp4).
#|   - Random mapping followed by coarse-graining. First frame, animation [available here](https://ninivert.github.io/lcnthesis/chapters/figures/coarse_graining_p=2_random.mp4).
#|   - Random mapping followed by coarse-graining. Last frame, animation [available here](https://ninivert.github.io/lcnthesis/chapters/figures/coarse_graining_p=2_random.mp4).
#|   - Column mapping followed by coarse-graining. First frame, animation [available here](https://ninivert.github.io/lcnthesis/chapters/figures/coarse_graining_p=2_column.mp4).
#|   - Column mapping followed by coarse-graining. Last frame, animation [available here](https://ninivert.github.io/lcnthesis/chapters/figures/coarse_graining_p=2_column.mp4).
#|   - Z-mapping followed by coarse-graining. First frame, animation [available here](https://ninivert.github.io/lcnthesis/chapters/figures/coarse_graining_p=2_z.mp4).
#|   - Z-mapping followed by coarse-graining. Last frame, animation [available here](https://ninivert.github.io/lcnthesis/chapters/figures/coarse_graining_p=2_z.mp4).
#| echo: false
# Note: I have to do this wizardry because Quarto *really* doesn't seem to like nested divs with conditional content...
from IPython.display import Image
display(Image('figures/coarse_graining_p=2_original.mp4-1.jpg'))
display(Image('figures/coarse_graining_p=2_original.mp4-2.jpg'))
display(Image('figures/coarse_graining_p=2_random.mp4-1.jpg'))
display(Image('figures/coarse_graining_p=2_random.mp4-2.jpg'))
display(Image('figures/coarse_graining_p=2_column.mp4-1.jpg'))
display(Image('figures/coarse_graining_p=2_column.mp4-2.jpg'))
display(Image('figures/coarse_graining_p=2_z.mp4-1.jpg'))
display(Image('figures/coarse_graining_p=2_z.mp4-2.jpg'))
```
:::

### Iterative application of $S$: from $[0,1]^3$ to $[0,1]^2$, then to $[0,1]$ {#sec-simulations-3d-2d-1d}

<!-- #### Continuity on one axis and fractal on the other: from $[0,1]^3$ to $[0,1]^2$ -->

In the previous section, we showed that the coarse-graining procedure works from $[0,1]^2$ to $[0,1]$. However, one might argue that the reason why it worked, is not the mapping, but rather the continuity of the neural field in $[0,1]^2$. To show the generality of our results, we show that coarse-graining works, even when the original neural field is fractal.

For this, we again consider the same cycling field, but with $p=3$, such that the natural embedding space is $\R^3$, with the equivalent formulation on $[0,1]^3$. We again consider a grid with a mesh size of $2^{-n}$ along each spatial dimension, resulting in $2^n \times 2^n \times 2^n = 8^n$ "cube populations", at positions $(v_1, v_2, v_3) \in [0,1]^3$.

To go from $[0,1]^3$ to $[0,1]^2$, we consider a slice $(v_2, v_3)$ at constant $v_1$, map $(v_2, v_3)$ to $\alpha_{23}$, then perform coarse-graining, which brings down the number of square populations in $[0,1]^2$ from $8^n$ to $4^n$. The result of this is an embedding of the neural field on $(v_1, \alpha_{23}) \in [0,1]^2$. The corresponding neural field is continuous in its $v_1$ component, but fractal in $\alpha_{23}$. This is visualized in @fig-cg-3d-2d-1d, where we see that in the state $h(z_1, z_2, z_3) = z_1$, the corresponding 2D neural field is constant along the $\alpha_{23}$ component, since the slice $(v_2, v_3)$ has a constant activity level. However, as soon as $h$ evolves to a new state, non-orthogonal to $z_2$ and $z_3$, the fractal structure becomes clear, because the slice $(v_2, v_3)$ is no longer constant.

::: {.content-visible when-format="html"}
![Equivalence of dynamics, from 3D to 2D to 1D.](figures/embedding_3d_2d_1d.mp4){#fig-cg-3d-2d-1d}
:::

::: {.content-visible when-format="pdf"}
```{python}
#| label: fig-cg-3d-2d-1d
#| layout-nrow: 2
#| fig-cap: Equivalence of dynamics, from 3D to 2D to 1D. The first and the last frame of the animation are shown, [available here](https://ninivert.github.io/lcnthesis/chapters/figures/embedding_3d_2d_1d.mp4).
#| fig-subcap:
#|   - First frame
#|   - Last frame
#| echo: false
# Note: I have to do this wizardry because Quarto *really* doesn't seem to like nested divs with conditional content...
from IPython.display import Image
display(Image('figures/embedding_3d_2d_1d.mp4-1.jpg'))
display(Image('figures/embedding_3d_2d_1d.mp4-2.jpg'))
```
:::

We close this chapter by repeating this process to go from $[0,1]^2$ to $[0,1]$, mapping $(v_1, \alpha_{23})$ to $\alpha_{123}$, and coarse-graining. In total, we have coarse-grained two times: once from $[0,1]^3$ ($8^n$ cube populations) to $[0,1]^2$ ($4^n$ square populations), and again from $[0,1]^2$ to $[0,1]$ ($2^n$ segment populations). The corresponding 1D neural field is now fully fractal, and again exhibits the same dynamics. This shows that "local mappings conserve regularity": we started with a fully continous neural field in $[0,1]^3$, and obtained a partially fractal neural field in $[0,1]^2$. Despite the fractal structure, the 2D neural field is still regular enough that the dynamics are conserved, but more importantly that it can *again* be mapped to an equivalent 1D neural field.
<!-- In @fig-averaging-geom-3d, we repeat a similar geometric view as @fig-averaging-geom: we see the Z-mapping in the slice $(v_2, v_3)$, how after coarse-graining the kernel is still effectively "sampled" at averaged locations, and crucially, how repeating the coarse-graining still leads to average locations that span the entire $[0,1]^3$ box (instead of being constrained to a plane embedded in $[0,1]^3$, as would have been the case with the Column mapping). -->

<!-- ![(TODO what $n$) Effective sampling points after averaging.](figures/mapping_3d_2d_1d.png){width=400px #fig-averaging-geom-3d} -->